---
title: "Data Prep"
author: "Olivia"
date: "2024-08-05"
output: html_document
---

##load libraries, data

```{r}
rm(list = ls())
library(tidyverse)
library(lubridate)
library(haven)
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(DiagrammeR)
library(DiagrammeRsvg)
library(rsvg)
library(glue)
library(gtsummary)
library(cardx)

setwd("C:/Users/mwhite10/Downloads")
data <- read_sas("./pubsaf2403/cand_kipa.sas7bdat")
stathist = read_sas("./pubsaf2403/stathist_kipa.sas7bdat")
ADI <- read_csv("US_2022_ADI_Census_Block_Group_v4_0_1.csv")
SVI <- read_csv("SVI_2022_US.csv")
SVM <- read_csv("SVM_ZIPCODE_SCORES_All States_All Counties_2024-06-24.csv")
DCI <- read_xlsx("dci-2017-2021-scores-only (2).xlsx")
zip_crosswalk <- read_excel("ZIP_TRACT_032024 (1).xlsx")
cand_zip <- read_sas("canzip2403.sas7bdat")
```

##set cohort - start and end dates, kidney (+/- pancreas) only, adults only

```{r}
start_date <- as.Date('2014-01-01')
end_date <- as.Date('2022-12-31')
filt_date <- data %>% filter(between(CAN_LISTING_DT, start_date, end_date))
filt_date_KIKP <- filt_date %>% filter(WL_ORG %in% c("KI", "KP"))
filt_date_KIKP %>% filter(CAN_AGE_AT_LISTING < 18) ##pediatric candidates
filt_date_KIKP_adult <- filt_date_KIKP %>% filter(CAN_AGE_AT_LISTING >= 18)

rm(filt_date, filt_date_KIKP, data)
```

##from here down based on Kevin Zhang's code
##establish waitlist_end_date - transplant date, removal date, last inactive date (if after last active date), last active date

```{r}
df_cand_kipa <- filt_date_KIKP_adult %>% ##df_cand_kipa has same cohort, adds waitlist_end_date column
  mutate(waitlist_end_date = case_when(
    is.na(REC_TX_DT) == FALSE ~ REC_TX_DT, ##look at transplant date first
    is.na(CAN_REM_DT) == FALSE ~ CAN_REM_DT, ##then look at removal date
    is.na(CAN_LAST_INACT_STAT_DT) == FALSE & CAN_LAST_INACT_STAT_DT > CAN_LAST_ACT_STAT_DT ~ CAN_LAST_INACT_STAT_DT,##then inactive date if after last active date
    !is.na(CAN_LAST_ACT_STAT_DT) ~ CAN_LAST_ACT_STAT_DT, ##then last active
    is.na(CAN_LAST_ACT_STAT_DT) & !is.na(CAN_LAST_INACT_STAT_DT) ~ CAN_LAST_INACT_STAT_DT,
    TRUE ~ CAN_LAST_ACT_STAT_DT)) ##if all else fails, use last active date

rm(filt_date_KIKP_adult)

df_cand_kipa = df_cand_kipa %>%
  group_by(PERS_ID) %>%
  mutate(all_centers = paste(unique(CAN_LISTING_CTR_CD), collapse = ",")) %>% 
  ungroup()

```

##single registrations - new columns num_list, min_list_date, wait_time, outcome

```{r}
single_registrations <- df_cand_kipa %>%
  group_by(PERS_ID) %>% ##grouping by PERS_ID - ?
  mutate(num_list = n()) %>%
  filter(num_list == 1) %>%
  ungroup() %>%
  mutate(min_list_date = CAN_LISTING_DT,  
         wait_time = waitlist_end_date - min_list_date, ##using end dates established in last code chunk
         outcome = case_when(
           DON_TY == "C" ~ "DDKT",
           DON_TY == "L" ~ "LDKT",
           !is.na(CAN_REM_CD) ~ "removed/died",  ##do we want all other CAN_REM_CD codes here? or just 8 and 13 (like used to define the outcome below)
           TRUE ~ "censored"
         ))
```

##multiple registrations - new columns num_list, list_type (concurrent or sequential), num_tx; fills missing REC_TX_DT

```{r}
multiple_registrations <- df_cand_kipa %>%
  filter(!PX_ID %in% single_registrations$PX_ID) %>%
  group_by(PERS_ID) %>%
  arrange(PERS_ID, CAN_LISTING_DT) %>%
  mutate(num_list = n())

candidates_w_multiple_registrations <- multiple_registrations %>% pull(PERS_ID) %>% unique() %>% length() ##number of candidates with multiple registrations

avg_num_registrations_multiple <- mean(multiple_registrations %>% group_by(PERS_ID) %>% filter(row_number() == 1) %>% pull(num_list)) ##for candidates with multiple registrations, the average number of registrations

multiple_registrations <- multiple_registrations %>%
  mutate(list_type = case_when(
    CAN_LISTING_DT < lag(waitlist_end_date) ~ "concurrent",
    waitlist_end_date > lead(CAN_LISTING_DT) ~ "concurrent",
    TRUE ~ "sequential")) %>%
  mutate(REC_TX_DT = as.Date(REC_TX_DT)) %>%
  mutate(num_tx = length(unique(na.omit(REC_TX_DT)))) %>%
  fill(REC_TX_DT, .direction='downup') ##this line is causing issues with negative dial_time values
```

```{r}
multiple_registrations <- multiple_registrations[order(multiple_registrations$PERS_ID, multiple_registrations$waitlist_end_date), ] ##orders by PERS_ID then waitlist_end_date

##multiple registrations - number sequential transplants, correct concurrent listing transplant dates

multiple_registrations$transplant_num <- 1 ##new column transplant_num, setting all at 1 for now
for (i in 2:nrow(multiple_registrations)) {
  if (!is.na(multiple_registrations$PERS_ID[i-1]) && 
      !is.na(multiple_registrations$PERS_ID[i]) && 
      !is.na(multiple_registrations$REC_TX_DT[i-1]) && 
      !is.na(multiple_registrations$REC_TX_DT[i]) &&
      multiple_registrations$PERS_ID[i-1] == multiple_registrations$PERS_ID[i] &&
      multiple_registrations$REC_TX_DT[i-1] != multiple_registrations$REC_TX_DT[i]) {
    
    multiple_registrations$transplant_num[i] = multiple_registrations$transplant_num[i-1] + 1
  }
}

for (i in 2:nrow(multiple_registrations)) {
  if (!is.na(multiple_registrations$PERS_ID[i-1]) && 
      !is.na(multiple_registrations$PERS_ID[i]) && 
      multiple_registrations$PERS_ID[i-1] == multiple_registrations$PERS_ID[i] &&
      !is.na(multiple_registrations$transplant_num[i-1]) && 
      multiple_registrations$transplant_num[i-1] != multiple_registrations$transplant_num[i] &&
      multiple_registrations$transplant_num[i-1] != 1) {
    
    multiple_registrations$transplant_num[i] = multiple_registrations$transplant_num[i-1]
  }
}

multiple_registrations$transplant_num[multiple_registrations$list_type == 'sequential'] <- 0 ##when list_type is sequential, transplant_num is set to 0. I still don't understand this step either.

for(i in 1:(nrow(multiple_registrations)-1)) { ##loop from second to lat row
  if(multiple_registrations$PERS_ID[i] == multiple_registrations$PERS_ID[i+1] &
     multiple_registrations$list_type[i] == 'concurrent' & multiple_registrations$list_type[i+1] == 'concurrent' &
     !is.na(multiple_registrations$REC_TX_DT[i]) & !is.na(multiple_registrations$REC_TX_DT[i+1]) &
     multiple_registrations$REC_TX_DT[i] < multiple_registrations$REC_TX_DT[i+1] ) {
    
    multiple_registrations$REC_TX_DT[i] <- multiple_registrations$REC_TX_DT[i+1] ##if PERS_ID is the same as the NEXT row, both have concurrent listings, transplant dates for both are not NA, and transplant date is earlier than the next row's transplant date - update to next row's transplant date
  }}

sequential_lists <- multiple_registrations %>%
  filter(list_type == "sequential") %>%
  mutate(min_list_date = CAN_LISTING_DT,
         wait_time = waitlist_end_date - min_list_date,
         outcome = case_when(
           DON_TY == "C" ~ "DDKT",
           DON_TY == "L" ~ "LDKT",
           is.na(CAN_REM_CD) == FALSE ~ "removed/died", ##again, should this be all CAN_REM_CD? or just 8 and 13 (since we included the rest in censored)
           TRUE ~ "censored"
         ))  ##same thing as done above for single registrations

max_retransplants <- max(multiple_registrations$transplant_num) ##highest number of transplants for a single person

multiple_registrations <- multiple_registrations %>% ##for candidates with multiple registrations, adds columns min_list_date, wait_time
  group_by(PERS_ID, transplant_num) %>%
  mutate(min_list_date = min(CAN_LISTING_DT, na.rm=T),
         wait_time = waitlist_end_date - min_list_date)

multiple_registrations <- multiple_registrations %>% mutate

collapsed_concurrent_registrations <- NULL ##establishes object to store coming data
for(i in 1:max_retransplants) { ##loop from first row to max_transplants
  
  collapsed_concurrent_registrations <- rbind(collapsed_concurrent_registrations, 
        
  multiple_registrations %>%
    filter(list_type == "concurrent" & transplant_num == i) %>% ##filters for concurrent listings where translpant_num is equal to current value of i (?)
    mutate(DON_TY = ifelse(DON_TY == "", NA, DON_TY),
           last_wait_date = max(waitlist_end_date, na.rm = TRUE)) %>% ##if DON_TY is empty, fill with NA, set last_wait_date to waitlist_end_date (why?)
    fill(REC_TX_DT, .direction = "up") %>%
    fill(DON_TY, .direction = "up") %>%
    fill(DONOR_ID, .direction = "up") %>%
    fill(CAN_REM_CD, .direction = "up") %>% ##fills missing values upwards
    mutate(wait_time = case_when(
      is.na(REC_TX_DT) == FALSE & transplant_num != '0' ~ REC_TX_DT- min_list_date, ### Ignore non-transplanted rows
      TRUE ~ last_wait_date - min_list_date),
      outcome = case_when(
        DON_TY == "C" ~ "DDKT",
        DON_TY == "L" ~ "LDKT",
        is.na(CAN_REM_CD) == FALSE ~ "removed/died", ##same question - CAN_REM_CD should be all? or just 8, 13?
        TRUE ~ "censored")   ##does the same we already did for single registrations and sequential listings
    ) %>%
    select(-c(waitlist_end_date, CAN_LISTING_DT, CAN_REM_DT)) %>% ##drops columns waitlist_end_date, CAN_LISTING_DT, CAN_REM_DT (just for cleanliness?)
    filter(row_number() ==1) %>% ##for concurrent listings, keeps only first row 
    
    mutate(last_wait_date = case_when(
      REC_TX_DT < last_wait_date ~ REC_TX_DT,
      TRUE ~last_wait_date)))}
```

##recreate CAN_LISTING_DT - to fix issue of NA values later on

```{r}
collapsed_concurrent_registrations <- collapsed_concurrent_registrations %>% mutate(CAN_LISTING_DT = min_list_date)
```

##recombine separated data frames

```{r}
df_cand_kipa <- bind_rows(single_registrations %>% ungroup(), 
                          sequential_lists %>% ungroup(), 
                          collapsed_concurrent_registrations %>% ungroup()) ##puts data all back together

df_cand_kipa %>% filter(is.na(CAN_LISTING_DT))
```

##establish final data frames. 3 outcomes = received deceased donor kidney transplant (unique_transplant), died or removed due to deterioration (rem_death_deterioration), censored (censored_total) 

```{r}
unique_candidates <- df_cand_kipa %>% pull(PERS_ID) %>% n_distinct()

unique_listings <- df_cand_kipa %>% nrow()

unique_transplant <- df_cand_kipa %>% filter(outcome == "DDKT")

rem_death_deterioration <- df_cand_kipa %>% filter(outcome == "removed/died") %>% filter(CAN_REM_CD %in% c(8, 13)) ##fixes my previous questions by only including 8 and 13 from removed/died classification

censored_total <- df_cand_kipa %>% filter(outcome %in% c("removed/died", "censored", "LDKT")) %>% filter(!(CAN_REM_CD %in% c(8, 13)) | (outcome == "LDKT"))
```

##break down censored groups

```{r}
ldkt <- df_cand_kipa %>% filter(outcome == "LDKT")

rem_other_reason <- df_cand_kipa %>% filter(outcome %in% c("censored", "removed/died")) %>% filter(!(CAN_REM_CD %in% c(8,13)))

rm(collapsed_concurrent_registrations, sequential_lists, single_registrations, i)
```

##creates dialysis_time_at_list (in years), 0 if preemptively listed; preemptive means preemptive_listing = 1 and dialysis = 0
```{r}
df_cand_kipa <- df_cand_kipa %>% 
  mutate(
    dialysis_time_at_list = case_when(
      is.na(CAN_DIAL_DT) | as.Date(min_list_date) < as.Date(CAN_DIAL_DT) ~ 0,
      TRUE ~ as.numeric(difftime(min_list_date, CAN_DIAL_DT, units = 'weeks')) / 52.25),
    preemptive_listing = case_when(
      dialysis_time_at_list == 0 ~ 1,
      TRUE ~ 0),
    dialysis = case_when(
      preemptive_listing == 0 ~ 1,
      TRUE ~ 0),
    living_donor = ifelse(outcome == "LDKT", 1, 0)
    )


```


```{r}

df_cand_kipa <- df_cand_kipa %>%
  mutate(
    diabetes_cat = case_when(
      CAN_DIAB_TY == '2' | CAN_DIAB_TY == '3' | CAN_DIAB_TY == '4' | CAN_DIAB_TY == '5' ~ 1, ##1 represents presence of diabetes
      TRUE ~ 0)) ##only returns 179,427 rows - only patients that received transplant
```



```{r}

df_cand_kipa <- df_cand_kipa %>%
  mutate(race = case_when(
    CAN_RACE_SRTR == 'WHITE' & CAN_ETHNICITY_SRTR == 'NLATIN' ~ 'White',
    CAN_RACE_SRTR == 'BLACK' & CAN_ETHNICITY_SRTR == 'NLATIN' ~ 'Black_AA',
    CAN_RACE_SRTR == 'NATIVE' & CAN_ETHNICITY_SRTR == 'NLATIN'  ~ 'American_Indian',
    CAN_RACE_SRTR == 'ASIAN' & CAN_ETHNICITY_SRTR == 'NLATIN'  ~ 'Asian',
    CAN_RACE_SRTR == 'PACIFIC' & CAN_ETHNICITY_SRTR == 'NLATIN'  ~ 'Pacific_Islander',
    CAN_RACE_SRTR == 'MULTI' & CAN_ETHNICITY_SRTR == 'NLATIN' ~ 'Multi',
    CAN_RACE_SRTR == 'NA' & CAN_ETHNICITY_SRTR == 'NLATIN' ~ 'Unknown',
    CAN_RACE_SRTR == '2000' & CAN_ETHNICITY_SRTR == 'LATINO' ~ 'Hispanic',
    TRUE ~ 'Other')) %>%
  
  mutate(sex = case_when(
    CAN_GENDER == 'M' ~ 'Male',  ##creates sex variable
    CAN_GENDER == 'F' ~ 'Female'),
    
    diabetes = factor(diabetes_cat, levels=c(0,1)),
    dialysis = factor(dialysis, levels=c(0,1)),
    preemptive_listing = factor(preemptive_listing, levels=c(0,1)),
    previous_TX = factor(CAN_PREV_TX, levels=c(0,1)),
    age = as.numeric(CAN_AGE_AT_LISTING)
  ) %>%
  
  mutate(
    age_group = case_when(
      age <= 24 ~ '18-24',
      age > 24 & age <= 29 ~ '25-29',
      age > 29 & age <= 39 ~ '30-39',
      age > 39 & age <= 49 ~ '40-49',
      age > 49 & age <= 59 ~ '50-59',
      age > 59 & age <= 69 ~ '60-69', ##establishes age groups
      TRUE ~ '>70'))

df_cand_kipa$age_group <- factor(df_cand_kipa$age_group, levels=c('18-24', '25-29', '30-39',
                                                                  '40-49', '50-59', '60-69', '>70'))  ##makes age groups factor variables
df_cand_kipa$race <- factor(df_cand_kipa$race, levels=c('White', 'American_Indian', 
                                                        'Asian', 'Black_AA', 'Hispanic',
                                                        'Other', 'Pacific_Islander', 'Unknown')) ##makes race a factor variable
### Initial variable selection
df_cand_kipa <- df_cand_kipa %>%
  select(PX_ID, PERS_ID, living_donor, 
         outcome, list_type, waitlist_end_date, wait_time, min_list_date, last_wait_date,
         age, age_group, sex, race, diabetes, dialysis, dialysis_time_at_list,
         preemptive_listing, previous_TX,
         CAN_LISTING_DT, CAN_DIAL_DT, CAN_REM_CD, CAN_REM_DT, REC_TX_DT, 
         CAN_LAST_INACT_STAT_DT, CAN_LAST_ACT_STAT_DT, 
         CAN_DEATH_DT, PERS_SSA_DEATH_DT, PERS_OPTN_DEATH_DT, CAN_RACE_SRTR, CAN_AGE_AT_LISTING, CAN_ETHNICITY_SRTR, CAN_TIEBREAKER_DT, CAN_GENDER, CAN_ABO, CAN_LISTING_CTR_CD, all_centers, CAN_PRIMARY_PAY) ##only leaves in necessary variables


### A death indicator variable 
df_cand_kipa <- df_cand_kipa %>% 
  mutate(death = case_when(
    CAN_REM_CD == '8' | !is.na(CAN_DEATH_DT) | !is.na(PERS_SSA_DEATH_DT) | !is.na(PERS_OPTN_DEATH_DT) ~ 1,
    TRUE ~ 0))

# If transplanted, transplanted takes priority
df_cand_kipa$death[df_cand_kipa$outcome == '>20% KDPI DDKT' | df_cand_kipa$outcome == 'DDKT' |
                   df_cand_kipa$outcome == 'LDKT' |  df_cand_kipa$outcome == 'top 20% KDPI DDKT'] <- 0
```







##ADI

```{r ADI}

ADI$FIPS_11_digit <- substr(ADI$FIPS, 1,11) ##first 11 digits of FIPS column make new column


ADI <- left_join(ADI,
                 select(zip_crosswalk, TRACT, ZIP),
                 by = c("FIPS_11_digit" = "TRACT")) ##merging TRACT and ZIP from crosswalk data with ADI. matches FIPS_11_digit with TRACT.


check_numeric_ADI <- unique(ADI$ADI_NATRANK[!is.na(ADI$ADI_NATRANK) & !grepl("^\\d+$", ADI$ADI_NATRANK)]) ##to find non-numeric NATRANK values

ADI <- ADI %>%
  group_by(ZIP) %>%
  mutate(ADI_percentile = mean(as.numeric(ADI_NATRANK), na.rm = TRUE)) ##new column mean_ADI. ignores NA values

ADI <- distinct(ADI, ZIP, .keep_all = T) 


```

##SVI

```{r}


SVI <- left_join(SVI, 
                 select(zip_crosswalk, TRACT, ZIP), 
                 by = c("FIPS" = "TRACT")) ##join  TRACT and ZIP from SVI_crosswalk with SVI, matching FIPS to tract.

SVI <- SVI %>% 
  select(ZIP, RPL_THEMES) 

SVI <- SVI %>% 
  mutate(RPL_THEMES = case_when(RPL_THEMES < 0 ~ NA, .default = RPL_THEMES)) 

SVI <- SVI %>% 
  group_by(ZIP) %>% 
  mutate(SVI_percentile = mean(as.numeric(RPL_THEMES*100), na.rm = T),
         median_SVI = median(as.numeric(RPL_THEMES*100), na.rm = T))

SVI <- distinct(SVI, ZIP, .keep_all = T)
```

## SVM
why the second line? if ZIPCODE is already 5 digits?
```{r}
SVM$zipcode <- substr(SVM$ZIPCODE, start = 1, stop = 5) 
```

##DCI
```{r}
DCI <- DCI %>%
  mutate(zip = as.character(`Zip Code`), ##new column "zip"
         dci_percentile = `2017-2021 Final Distress Score`) ##new column dci_percentile which is exactly the distress score

DCI <- DCI %>%
  mutate(`Zip Code` = as.character(`Zip Code`)) %>%    ##gives 4-digit zips a leading zero
  mutate(`Zip Code` = if_else(str_length(`Zip Code`) == 4,
                              str_pad(`Zip Code`, width = 5, side = "left", pad = "0"),
                              `Zip Code`))
```

##Assign deprivation indicies to candidate zip code - credit to alex wang

```{r}

cand_zip$ZIP = substr(cand_zip$CAN_PERM_ZIP, start = 1, stop = 5) ##extracts first 5 digits from CAN_PERM_ZIP and creates new column ZIP

cand_zip <- cand_zip %>%
  left_join(select(SVM, ZIPCODE, SVM_SCORE, SVM_PERCENTILE),
                      by = c("ZIP" = "ZIPCODE")) ##join ZIPCODE, SVM_SCORE, SVM_PERCENTILE from SVM_ZIP with cand_zip. merge ZIP and ZIPCODE
cand_zip <- left_join(cand_zip, select(ADI, ADI_percentile, ZIP), by = c("ZIP" = "ZIP"))
cand_zip <- left_join(cand_zip, select(DCI, dci_percentile, zip), by = c("ZIP" = "zip"))
cand_zip <-left_join(cand_zip, SVI, by = c("ZIP" = "ZIP"))

```

```{r}
rm(ADI, ADI_crosswalk, DCI, SVI, SVI_crosswalk, SVM, check_numeric_ADI)
```

##start with our established cohort, add column for preemptive listing

```{r}
df_cand_kipa <- df_cand_kipa %>%
  mutate(preemp = case_when(
    CAN_DIAL_DT > CAN_LISTING_DT ~ TRUE,
    is.na(CAN_DIAL_DT) ~ TRUE,
    is.na(CAN_LISTING_DT) & CAN_DIAL_DT > CAN_TIEBREAKER_DT ~ TRUE,
    TRUE ~ FALSE
  ))

df_cand_kipa <- df_cand_kipa %>%
  mutate(dial_time = NA) %>%  # Initialize dial_time with NA
  mutate(dial_time = case_when(
    preemp == FALSE ~ CAN_LISTING_DT - CAN_DIAL_DT,
    TRUE ~ dial_time  # Keep existing dial_time for preemp == TRUE
  ))

```

```{r}
df_cand_kipa %>% select(PX_ID, preemp, CAN_LISTING_DT, CAN_TIEBREAKER_DT, CAN_DIAL_DT, dial_time) %>% filter(is.na(dial_time))
df_cand_kipa %>% select(PX_ID, preemp, CAN_LISTING_DT, CAN_TIEBREAKER_DT, CAN_DIAL_DT) %>% filter(is.na(preemp))
```



```{r}
df_cand_kipa_w_scores <- left_join(df_cand_kipa, select(cand_zip, SVM_SCORE, SVM_PERCENTILE, PX_ID, SVI_percentile, ADI_percentile, dci_percentile),
                              by = c("PX_ID" = "PX_ID"))

n_distinct(df_cand_kipa_w_scores$PX_ID)
```



```{r}
df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>%
  mutate(SVM_PERCENTILE = as.numeric(as.character(SVM_PERCENTILE)),
         SVI_percentile = as.numeric(as.character(SVI_percentile)),
         ADI_percentile = as.numeric(as.character(ADI_percentile)),
         dci_percentile = as.numeric(as.character(dci_percentile)))

df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>% 
  group_by(PX_ID) %>% ##why group by PX_ID? shouldn't they all be unique?
  mutate(SVM_decile = cut(SVM_PERCENTILE, c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10")),
                                                                    SVI_decile = cut(SVI_percentile, c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10")),
                                                                    ADI_decile = cut(ADI_percentile,c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10")),
                                                                    DCI_decile = cut(dci_percentile,c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10")))

df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>%
  mutate(svm_decile = as.numeric(SVM_decile),
        svi_decile = as.numeric(SVI_decile),
        adi_decile = as.numeric(ADI_decile),
        dci_decile = as.numeric(DCI_decile))

```

##df_cand_kipa_w_scores is now the data frame that will be used for major analyses.

##add a column with race as a factor 

```{r}
df_cand_kipa_w_scores$can_race_factor <- as.factor(df_cand_kipa_w_scores$CAN_RACE_SRTR)
df_cand_kipa_w_scores$SVM_decile <- as.factor(df_cand_kipa_w_scores$SVM_decile)


df_cand_kipa_w_scores$can_race_factor <- relevel(df_cand_kipa_w_scores$can_race_factor, ref = "WHITE")
df_cand_kipa_w_scores$SVM_decile <- relevel(df_cand_kipa_w_scores$SVM_decile, ref = "5")
```

##establish columns start_date, transplant_date, removal_date, last_active_date, last_inactive_date, death_date, death_date_max, transplant_time, death_time, removal_time

```{r}
df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>%
  mutate(
    start_date = as.Date(ifelse(is.na(CAN_LISTING_DT), CAN_TIEBREAKER_DT, CAN_LISTING_DT)), ##added by edgar, takes care of 33,093 survival_time == NA values
    transplant_date = as.Date(REC_TX_DT),
    removal_date = as.Date(CAN_REM_DT),
    last_active_date = as.Date(CAN_LAST_ACT_STAT_DT),
    last_inactive_date = as.Date(CAN_LAST_INACT_STAT_DT),
    death_date = as.Date(CAN_DEATH_DT),
    death_date_max = pmax(CAN_DEATH_DT, PERS_SSA_DEATH_DT, PERS_OPTN_DEATH_DT, na.rm = T)) %>% ##do we want max of these? or is PERS_OPTN_DEATH_DT most complete?
  mutate(transplant_time = ifelse(!is.na(transplant_date),
      as.numeric(transplant_date - start_date, units = 'days'), NA)) %>%
  mutate(death_time = ifelse(!is.na(death_date_max),
      as.numeric(death_date_max - start_date, units = 'days'), NA)) %>%
  mutate(removal_time = as.numeric(
      pmax(removal_date, last_active_date, last_inactive_date, na.rm = T) - start_date,
      units = 'days'))

n_distinct(df_cand_kipa_w_scores$PX_ID)
```

##add column survival_time with the minimum of transplant, death, and removal time. cap survival time at 5 years

```{r}
df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>%
  mutate(
    transplant_time = ifelse(transplant_time < 0, NA, transplant_time),
    death_time = ifelse(death_time < 0, NA, death_time),
    removal_time = ifelse(removal_time < 0, NA, removal_time),
    survival_time = pmin(transplant_time, death_time, removal_time, na.rm = TRUE),
    survival_time = case_when(
      survival_time > 1826 ~ 1826,
      .default = survival_time ## I noticed if one of the 3 outcomes was negative but another was positive in some cases it was using the negative one. this should take care of that? Took us from 70 negative survival_times to 0...but back up from 0 to 8 survival_times == NA. this is because if a survival_time was going to be negative it replaced it with NA
    )
  )

```

##new column classifiying outcomes - numbers match from when we defined cohort except one extra in "transplant"

```{r}
df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>% 
  mutate(status = case_when(
    outcome == "DDKT" & survival_time < 1826 ~ 1,  # Transplanted
    outcome == "removed/died" & CAN_REM_CD == 8 & survival_time < 1826 ~ 2,  # Removed/died due to deterioration
    outcome == "LDKT" & survival_time < 1826 ~ 3,  #censored
    outcome == "removed/died" & CAN_REM_CD != 8 & survival_time < 1826 ~ 4,
    TRUE ~ 0  # Default to 0 for any other cases
  ))

df_cand_kipa_w_scores$status <- factor(df_cand_kipa_w_scores$status,
                                       levels = c(0,1,2,3,4),
                                       labels = c("censor", "DDKT", "death", "LDKT", "removed"))

table(df_cand_kipa_w_scores$status)
```

##add transplant_outcome and death_outcome columns

```{r}
df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>%
  mutate(transplant_outcome = case_when(
    survival_time == 1826 ~ 0,
    status == "censor" ~ 0,
    status == "DDKT" ~ 1,
    .default = 0),
    death_outcome = case_when(
      survival_time == 1826 & !(CAN_REM_CD %in% c(8,13)) ~ 0,
      status == "censor" ~ 0,
      status == "death" ~ 1,
      .default = 0
    ))
```

##add sex , latino, status_numeric columns (?)

```{r}
df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>%
  mutate(sex = case_when(
    CAN_GENDER == "M" ~ 1,
    CAN_GENDER == "F" ~ 0),
    latino = case_when(
      CAN_ETHNICITY_SRTR == "LATINO" ~1,
      CAN_ETHNICITY_SRTR == "NLATIN" ~0,
      .default = NA),
    status_numeric = case_when(
      status == "censor" ~ 0,
      status == "DDKT" ~ 1,
      status == "death" ~ 2,
      status == "LDKT" ~ 3,
      status == "removed" ~ 4
    ))
```

```{r}

#there was a duplicated row, take out
df_cand_kipa_w_scores = df_cand_kipa_w_scores %>% distinct()
table(df_cand_kipa_w_scores$status)
```


```{r}


strobe = grViz(
  glue("digraph my_flowchart {{ 
      graph[splines = ortho]
      node [fontname = Helvetica, shape = box, width = 4, height = 1]
      
        // Main node
        node1[label = <Kidney Transplant Candidate Registrations, 2014-2022 (n = 362,752)>]
                
        // Invisible blank node to adjust positioning
        blank1[label = '', width = 0.01, height = 0.01]
        
        // Excluded criteria (shift to left)
        excluded1[label = <Pediatric Candidates (n = 9,513)>]
        
        // Connect Total available patients to blank1 (no arrow head)
        node1 -> blank1[dir = none];
        
        // Connect blank1 to excluded1 (on the left side)
        excluded1 -> blank1[dir = back, minlen = 2];
        
        // Keep excluded1 on the left of blank1
        {{ rank = same; blank1 excluded1 }}

        // Blank node for the next exclusion step
        blank2[label = '', width = 0.01, height = 0.01]
        
        // Excluded due to missing values (same rank)
        excluded2[label = <Multiple Listings (n = 40,272)>]
        
        // Connect blank1 to blank2 (no arrow head)
        blank1 -> blank2[dir = none];
        
        // Connect blank2 to excluded2 (right side)
        blank2 -> excluded2[minlen = 2];
        
        // Keep blank2 and excluded2 aligned in the same rank
        {{ rank = same; blank2 excluded2 }}
        
        // Connect blank2 to next step (included for analysis)
        node2[label = <Adult Kidney Transplant Candidate Registrations (n = 312,967)>]
        blank2 -> node2;
        
        // Connect to data linking
        node3[label = <Received Kidney Transplant (DDKT) (n = 114,103)>]
        node4[label = <Died (n = 25,503)>]
        node5[label = <Censored (n = 68,533)>]
        node6[label = <Living Donor kidney transplant (n = 48,289)>]
        node7[label = <Removed for Other Reason (n = 56,538)>]
        
        node2 -> node3;
        node2 -> node4;
        node2 -> node5;
        node2 -> node6;
        node2 -> node7;
     }}")
)

diagram = export_svg(strobe)
strobe
rsvg_png(charToRaw(diagram), "strobe_diagram.png",  width = 10000, height = 3000)
```




```{r}

df_cand_kipa_w_scores = df_cand_kipa_w_scores %>% mutate(insurance = case_when(
    CAN_PRIMARY_PAY %in% c(1) ~ "Private insurance",
    CAN_PRIMARY_PAY %in% c(2, 3, 4, 5, 6, 7) ~ "Public insurance",
    CAN_PRIMARY_PAY %in% c(8) ~ "Self",
    CAN_PRIMARY_PAY %in% c(9) ~ "Donation",
    CAN_PRIMARY_PAY %in% c(10) ~ "Free Care",
    CAN_PRIMARY_PAY %in% c(11) ~ "Pending",
    CAN_PRIMARY_PAY %in% c(12) ~ "Foreign Government",
    CAN_PRIMARY_PAY %in% c(13) ~ "Medicare Unspecified",
    CAN_PRIMARY_PAY %in% c(14) ~ "US/State Govt Agency",
    CAN_PRIMARY_PAY %in% c(15) ~ "Unknown",
    TRUE ~ "Other"))
```

## Table 1 comparing patients with missing index scores to patients with full information

```{r}

missing_summary <- function(data,  label) {
  data %>%
    droplevels() %>%
    tbl_summary(
      include = c(CAN_AGE_AT_LISTING, CAN_GENDER, CAN_RACE_SRTR, CAN_ETHNICITY_SRTR,  preemptive_listing, dial_time, insurance),

      missing_text = "(Missing)"
    ) %>%
    modify_header(label = glue::glue("**{label}**")) %>%  # Dynamic header label                                # Bold significant p-values
    bold_labels()                                         # Bold variable labels
}

# Creating summaries for each index
svi_summary <- missing_summary(df_cand_kipa_w_scores %>% filter(is.na(SVI_decile)), "SVI")
adi_summary <- missing_summary(df_cand_kipa_w_scores %>% filter(is.na(ADI_decile)), "ADI")
svm_summary <- missing_summary(df_cand_kipa_w_scores %>% filter(is.na(SVM_decile)), "SVM")
dci_summary <- missing_summary(df_cand_kipa_w_scores %>% filter(is.na(DCI_decile)), "DCI")
non_missing_summary <- missing_summary(df_cand_kipa_w_scores %>% filter(!is.na(SVI_decile) & !is.na(ADI_decile) & !is.na(SVM_decile) & !is.na(DCI_decile)), "No Missing")


# Combining all tables into one
combined_summary <- tbl_merge(
  list(non_missing_summary, svi_summary, adi_summary, svm_summary, dci_summary),
  tab_spanner = c("**No Missing**", "**SVI**", "**ADI**", "**SVM**", "**DCI**")
)

# Display the combined summary
combined_summary

df_cand_kipa_w_scores %>% filter(!is.na(SVI_decile) & !is.na(ADI_decile) & !is.na(SVM_decile) & !is.na(DCI_decile))
```









```{r}
# Adjust 0 length survival times
df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>% mutate(
  survival_time = ifelse(survival_time == 0, 1, survival_time)
)
```



```{r}
df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>%
  mutate(SVM_PERCENTILE = as.numeric(as.character(SVM_PERCENTILE)), # SVM
         SVI_percentile = as.numeric(as.character(SVI_percentile)), # SVI
         ADI_percentile = as.numeric(as.character(ADI_percentile)), # ADI
         dci_percentile = as.numeric(as.character(dci_percentile))) # DCI

# Use median score from transplant center listed at for candidates missing score percentile
df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>%
  group_by(CAN_LISTING_CTR_CD) %>% # Gather by transplant center
  mutate(median_SVM = median(SVM_PERCENTILE, na.rm = T), # Median SVM of each transplant center
         median_SVI = median(SVI_percentile, na.rm = T), # Median SVI of each transplant center
         median_ADI = median(ADI_percentile, na.rm = T), # Median ADI of each transplant center
         median_DCI = median(dci_percentile, na.rm = T), # Median DCI of each transplant center
         SVM_PERCENTILE = ifelse(is.na(SVM_PERCENTILE), median_SVM, SVM_PERCENTILE), #conditional to input corresponding composite score value
         SVI_percentile = ifelse(is.na(SVI_percentile), median_SVI, SVI_percentile),  #conditional to input corresponding composite score value
         ADI_percentile = ifelse(is.na(ADI_percentile), median_ADI, ADI_percentile),  #conditional to input corresponding composite score value
         dci_percentile = ifelse(is.na(dci_percentile), median_DCI, dci_percentile))  #conditional to input corresponding composite score value
# Split Percentiles into Deciles
df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>% 
  group_by(PX_ID) %>% 
  mutate(SVM_decile = 
           cut(SVM_PERCENTILE, c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10")),
           SVI_decile = cut(SVI_percentile, c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10")),
           ADI_decile = cut(ADI_percentile,c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10")),
         DCI_decile = cut(dci_percentile,c(-Inf, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100), labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10")))

# Turn deciles into factors
df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>% 
  mutate(SVM_decile = as.factor(SVM_decile),
        SVI_decile = as.factor(SVI_decile),
        ADI_decile = as.factor(ADI_decile),
        DCI_decile = as.factor(DCI_decile))
```

```{r}
# blood type ) variable
df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>%
  mutate(typeO = ifelse(CAN_ABO == "O", 1, 0))

```


```{r}

# filter stathist to patients in dataframe
stathist_list = stathist %>% filter(PX_ID %in% df_cand_kipa_w_scores$PX_ID & CAN_LISTING_DT == CANHX_BEGIN_DT)

#recover cPRA from stathist
df_cand_kipa_w_scores = df_cand_kipa_w_scores %>% left_join(stathist_list, select('CANHX_CPRA'), by = "PX_ID")

#set missing values to 0
df_cand_kipa_w_scores = df_cand_kipa_w_scores %>% mutate(
  CPRA = ifelse(!is.na(CANHX_CPRA), 100* CANHX_CPRA, 0)
)

#establish cpra score based on policy cutoffs
df_cand_kipa_w_scores = df_cand_kipa_w_scores %>% mutate(
  cpra_score = case_when(
    CPRA < 20 ~ "CPRA 0-20",
    CPRA >= 20 & CPRA < 30 ~ "CPRA 20-29",
    CPRA >= 30 & CPRA < 40 ~ "CPRA 30-39",
    CPRA >= 40 & CPRA < 50 ~ "CPRA 40-49",
    CPRA >= 50 & CPRA < 60 ~ "CPRA 50-59",
    CPRA >= 60 & CPRA < 70 ~ "CPRA 60-69",
    CPRA >= 70 & CPRA < 75 ~ "CPRA 70-74",
    CPRA >= 75 & CPRA < 80 ~ "CPRA 75-79",
    CPRA >= 80 & CPRA < 85 ~ "CPRA 80-84",
    CPRA >= 85 & CPRA < 90 ~ "CPRA 85-89",
    CPRA > 90 & CPRA < 95 ~ "CPRA 90-94",
    CPRA >= 95 & CPRA < 96 ~ "CPRA 95",
    CPRA >= 96 & CPRA < 97 ~ "CPRA 96", 
    CPRA >= 97 & CPRA < 98 ~ "CPRA 97",
    CPRA >= 98 & CPRA < 99 ~ "CPRA 98",
    CPRA >= 99 & CPRA < 100 ~ "CPRA 99",
    CPRA >= 100 ~ "CPRA 100"
  )
)
df_cand_kipa_w_scores = df_cand_kipa_w_scores %>% mutate(
  ldkt_outcome = ifelse(status == "LDKT", 1, 0)
)

```


```{r}
#create time variable


df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time))


```


```{r}
#bin dialysis time
df_cand_kipa_w_scores = df_cand_kipa_w_scores %>% mutate(
  dial_time = ifelse(is.na(dial_time), 0, dial_time)
) %>% mutate(
  dial_time_bins = case_when(
    dial_time == 0 ~ "Preemptive",
    dial_time > 0 & dial_time <= 365 ~ "Less than 1 year", 
    dial_time > 365 & dial_time <= 1095 ~ "1 to 3 years",
    dial_time > 1095 ~ "Greater than 3 years"
  )
)

```




```{r}
save(df_cand_kipa_w_scores, file = "df_cand_kipa_w_scores.RData")

```



