---
title: "dialysis_time_association"
author: "Edgar Aguila"
output: html_notebook
---

# Load in Libraries
```{r}
library(ggplot2)
library(tidyverse)
library(lubridate)
library(haven)
library(dplyr)
```

# Set reference levels to 5th decile
```{r}
study_cohort_df$SVM_decile <- relevel(study_cohort_df$SVM_decile, ref = "5")
study_cohort_df$SVI_decile <- relevel(study_cohort_df$SVI_decile, ref = "5")
study_cohort_df$ADI_decile <- relevel(study_cohort_df$ADI_decile, ref = "5")
study_cohort_df$DCI_decile <- relevel(study_cohort_df$DCI_decile, ref = "5")
```

## Linear Model w/ Dialysis Time
```{r}
# Create models
## SVM
svm_lm <- lm(dialysis_time_at_list ~ SVM_decile, data = study_cohort_df)

## SVI
svi_lm <- lm(dialysis_time_at_list ~ SVI_decile, data = study_cohort_df)

## ADI
adi_lm <- lm(dialysis_time_at_list ~ ADI_decile, data = study_cohort_df)

## DCI
dci_lm <- lm(dialysis_time_at_list ~ DCI_decile, data = study_cohort_df)

# View all regressions
svm_lm
svi_lm
adi_lm
dci_lm
```

# Linear Model Graph
```{r}
# Create new data frame
decile <- c(1:4, 6:10)

lm_df <- data.frame(decile = decile) # Convert to data frame
lm_df # view

# Create Coefficient columns
lm_df <- lm_df %>%
  mutate(
    exp_coeff_svm = exp(coef(svm_lm)[-1]), # SVM
    exp_coeff_svi = exp(coef(svi_lm)[-1]), # SVI
    exp_coeff_adi = exp(coef(adi_lm)[-1]), # ADI
    exp_coeff_dci = exp(coef(dci_lm)[-1]) # DCI
  )
lm_df # view

# Insert baseline row
lm_df <- lm_df %>%
  rbind(data.frame(decile = 5, exp_coeff_svm = 1, exp_coeff_svi = 1, exp_coeff_adi = 1, exp_coeff_dci = 1))

lm_df <- lm_df[order(lm_df$decile), ] # Put decile column in order
lm_df # view

# Create Graph
lm_df_visual <- ggplot(lm_df, aes(x = decile)) + # Set base
  geom_line(aes(y = exp_coeff_svm, color = "SVM")) + # SVM
  geom_line(aes(y = exp_coeff_svi, color = "SVI")) + # SVI
  geom_line(aes(y = exp_coeff_adi, color = "ADI")) + # ADI
  geom_line(aes(y = exp_coeff_dci, color = "DCI")) + # DCI
  theme_classic() + # Set theme (preference)
  labs(title = "Linear Model Coefficients of Dialysis by Decile", # Set title
       subtitle = "Data from SRTR 2024", # Set subtitle
       x = "Decile", # Set x-axis label
       y = "Correlation Coefficient", # Set y-axis label
       color = "Index") + # Set legend title
  scale_x_continuous(breaks = 1:10, # Create tick markers
                     labels = 1:10)
lm_df_visual # view

# Save to PDF
ggsave("/Users/edgaraguila/Desktop/School/Research Projects/UChicago Research/Data/dialysis_time_association Graphs/Linear Model Graph.pdf", plot = lm_df_visual, height = 4, units = "in")
```

# Log transformation of linear models
```{r}
# Log Transformation
study_cohort_df <- study_cohort_df %>%
  mutate(log_dialysis = case_when(
    dialysis_time_at_list > 0 ~ log(dialysis_time_at_list), # greater than zero, keep
    dialysis_time_at_list < 0 ~ log(dialysis_time_at_list), # less than zero, keep
    dialysis_time_at_list == 0 ~ log(1) # equal to zero, set log(1) so log value is 0
  ))
study_cohort_df

## Create Log Models
## SVM
svm_lm_log <- lm(log_dialysis ~ SVM_decile, data = study_cohort_df)

## SVI
svi_lm_log <- lm(log_dialysis ~ SVI_decile, data = study_cohort_df)

## ADI
adi_lm_log <- lm(log_dialysis ~ ADI_decile, data = study_cohort_df)

## DCI
dci_lm_log <- lm(log_dialysis ~ DCI_decile, data = study_cohort_df)

# View all regressions
svm_lm_log
svi_lm_log
adi_lm_log
dci_lm_log
```

# Log(Linear Model) Graph
```{r}
# Create new data frame
decile <- c(1:4, 6:10)

lm_log_df <- data.frame(decile = decile) # Convert to data frame
lm_log_df # view

# Create Coefficient columns
lm_log_df <- lm_log_df %>%
  mutate(
    exp_coeff_svm = exp(coef(svm_lm_log)[-1]), # SVM
    exp_coeff_svi = exp(coef(svi_lm_log)[-1]), # SVI
    exp_coeff_adi = exp(coef(adi_lm_log)[-1]), # ADI
    exp_coeff_dci = exp(coef(dci_lm_log)[-1]) # DCI
  )
lm_log_df # view

# Insert baseline row
lm_log_df <- lm_log_df %>%
  rbind(data.frame(decile = 5, exp_coeff_svm = 1, exp_coeff_svi = 1, exp_coeff_adi = 1, exp_coeff_dci = 1))

lm_log_df <- lm_log_df[order(lm_log_df$decile), ] # Put decile column in order
lm_log_df # view

# Create Graph
lm_log_df_visual <- ggplot(lm_log_df, aes(x = decile)) + # Set base
  geom_line(aes(y = exp_coeff_svm, color = "SVM")) + # SVM
  geom_line(aes(y = exp_coeff_svi, color = "SVI")) + # SVI
  geom_line(aes(y = exp_coeff_adi, color = "ADI")) + # ADI
  geom_line(aes(y = exp_coeff_dci, color = "DCI")) + # DCI
  theme_classic() + # Set theme (preference)
  labs(title = "Linear Model Coefficients of Log of Dialysis by Decile", # Set title
       subtitle = "Data from SRTR 2024", # Set subtitle
       x = "Decile", # Set x-axis label
       y = "Correlation Coefficient Log of Dialysis", # Set y-axis label
       color = "Index") + # Set legend title
  scale_x_continuous(breaks = 1:10, # Create tick markers
                     labels = 1:10)
lm_log_df_visual # view

# Save to PDF
ggsave("/Users/edgaraguila/Desktop/School/Research Projects/UChicago Research/Data/dialysis_time_association Graphs/Log of Linear Model Graph.pdf", plot = lm_log_df_visual, height = 4, units = "in")
```

# Preemptive Listing v. Decile (Logistic Regression)
```{r}
# Create models
## SVM
svm_glm <- glm(preemptive_listing ~ SVM_decile, data = study_cohort_df, family = "binomial")

## SVI
svi_glm <- glm(preemptive_listing ~ SVI_decile, data = study_cohort_df, family = "binomial")

## ADI
adi_glm <- glm(preemptive_listing ~ ADI_decile, data = study_cohort_df, family = "binomial")

## DCI
dci_glm <- glm(preemptive_listing ~ DCI_decile, data = study_cohort_df, family = "binomial")

# View all logistic regressions
svm_glm
svi_glm
adi_glm
dci_glm
```

# Logistic Regression Graph
```{r}
# Create new data frame
decile <- c(1:4, 6:10)

lr_df <- data.frame(decile = decile) # Convert to data frame
lr_df # view

# Create Coefficient columns
lr_df <- lr_df %>%
  mutate(
    SVM = exp(coef(svm_glm)[-1]), # SVM
    SVI = exp(coef(svi_glm)[-1]), # SVI
    ADI = exp(coef(adi_glm)[-1]), # ADI
    DCI = exp(coef(dci_glm)[-1]) # DCI
  )
lr_df # view

# Insert baseline row
lr_df <- lr_df %>%
  rbind(data.frame(decile = 5, SVM = 1, SVI = 1, ADI = 1, DCI = 1))

lr_df <- lr_df[order(lr_df$decile), ] # Put decile column in order
lr_df # view

# Create Graph
lr_df_visual <- ggplot(lr_df, aes(x = decile)) + # Set base
  geom_line(aes(y = SVM, color = "SVM")) + # SVM
  geom_line(aes(y = SVI, color = "SVI")) + # SVI
  geom_line(aes(y = ADI, color = "ADI")) + # ADI
  geom_line(aes(y = DCI, color = "DCI")) + # DCI
  theme_classic() + # Set theme (preference)
  labs(title = "Logistic Regression Coefficients of Preemptive Listing by Decile", # Set title
       subtitle = "Data from SRTR 2024", # Set subtitle
       x = "Decile", # Set x-axis label
       y = "Correlation Coefficient of Preemptive Listing", # Set y-axis label
       color = "Index") + # Set legend title
  scale_x_continuous(breaks = 1:10, # Create tick markers
                     labels = 1:10)
lr_df_visual # view

# Save to PDF
ggsave("/Users/edgaraguila/Desktop/School/Research Projects/UChicago Research/Data/dialysis_time_association Graphs/Logistic Regression of Preemptive Listing Graph.pdf", plot = lr_df_visual, height = 4, units = "in")
```

# Relevel factors to 1 (ask molly about this, there has to be an easier way)
```{r}
study_cohort_df$SVM_decile <- relevel(study_cohort_df$SVM_decile, ref = "4")
study_cohort_df$SVM_decile <- relevel(study_cohort_df$SVM_decile, ref = "3")
study_cohort_df$SVM_decile <- relevel(study_cohort_df$SVM_decile, ref = "2")
study_cohort_df$SVM_decile <- relevel(study_cohort_df$SVM_decile, ref = "1")

study_cohort_df$SVI_decile <- relevel(study_cohort_df$SVI_decile, ref = "4")
study_cohort_df$SVI_decile <- relevel(study_cohort_df$SVI_decile, ref = "3")
study_cohort_df$SVI_decile <- relevel(study_cohort_df$SVI_decile, ref = "2")
study_cohort_df$SVI_decile <- relevel(study_cohort_df$SVI_decile, ref = "1")

study_cohort_df$ADI_decile <- relevel(study_cohort_df$ADI_decile, ref = "4")
study_cohort_df$ADI_decile <- relevel(study_cohort_df$ADI_decile, ref = "3")
study_cohort_df$ADI_decile <- relevel(study_cohort_df$ADI_decile, ref = "2")
study_cohort_df$ADI_decile <- relevel(study_cohort_df$ADI_decile, ref = "1")

study_cohort_df$DCI_decile <- relevel(study_cohort_df$DCI_decile, ref = "4")
study_cohort_df$DCI_decile <- relevel(study_cohort_df$DCI_decile, ref = "3")
study_cohort_df$DCI_decile <- relevel(study_cohort_df$DCI_decile, ref = "2")
study_cohort_df$DCI_decile <- relevel(study_cohort_df$DCI_decile, ref = "1")
```

# Median Dialysis Time v. Decile Graph (Working)
```{r}
# Median Dialysis Time v. Decile (visual)
# Convert dialysis_time_at_list back to years
study_cohort_df <- study_cohort_df %>%
  mutate(dialysis_time_at_list = case_when(
    is.na(CAN_DIAL_DT) | as.Date(min_list_date) < as.Date(CAN_DIAL_DT) ~ 0, # Input 0 for patients never on dialysis
    T ~ as.numeric(difftime(min_list_date, CAN_DIAL_DT)) / 365 # Turn to year
  ))

# SVM
svm_dialysis <- study_cohort_df %>% # Create data frame for SVM Decile v. Dialysis Time
  group_by(SVM_decile) %>% # Group by decile
  summarize(SVM = dialysis_time_at_list) # Calculate median
svm_dialysis # view

# SVI
svi_dialysis <- study_cohort_df %>% # Create data frame for SVI Decile v. Dialysis Time
  group_by(SVI_decile) %>% # Group by decile
  summarize(SVI = dialysis_time_at_list) # Calculate median
svi_dialysis # view

# ADI
adi_dialysis <- study_cohort_df %>% # Create data frame for ADI Decile v. Dialysis Time
  group_by(ADI_decile) %>% # Group by decile
  summarize(ADI = dialysis_time_at_list) # Calculate median
adi_dialysis # view

# DCI
dci_dialysis <- study_cohort_df %>% # Create data frame for DCI Decile v. Dialysis Time
  group_by(DCI_decile) %>% # Group by decile
  summarize(DCI = dialysis_time_at_list) # Calculate median
dci_dialysis # view

# Join data frames
dialysis_v_decile <- bind_cols(svm_dialysis, svi_dialysis, adi_dialysis, dci_dialysis) %>% # Combine all data
  mutate(decile = SVM_decile) %>% # Change one decile column to "decile"
  select(-SVM_decile, -SVI_decile, -ADI_decile, -DCI_decile) # Remove other decile columns
dialysis_v_decile

rm(svm_dialysis, svi_dialysis, adi_dialysis, dci_dialysis) # Remove intermediate data frames

# Reshape data frame
dialysis_v_decile <- dialysis_v_decile %>%
  pivot_longer(cols = c(SVM, SVI, ADI, DCI), # Select columns
               names_to = "index", # Name of columns to "index"
               values_to = "dialysis_time") # values of column to "dialysis_time

dialysis_v_decile$decile <- as.factor(dialysis_v_decile$decile) # Set decile to factor

# Visualize graph
dialysis_v_decile_viz <- dialysis_v_decile %>% # Use this data frame
  ggplot(aes(x = decile, y = dialysis_time, fill = index)) + # Set x-axis
  geom_boxplot() + # Create Boxplot for graph
  facet_wrap(~ index) + # Create grid
  theme_bw() + # Set theme to classic (preference)
  scale_y_continuous(limits = c(0, 5), breaks = seq(0, 5, by = 1)) + # Create specific y-axis scale
  labs(title = "Dialysis Time at Listing by Decile", # Set title
       subtitle = "Data from SRTR 2024", # Set subtitle
       x = "Decile", # Set x-axis label
       y = "Dialysis Time at Listing (Year)", # Set y-axis label
       fill = "Index") # Set legend title
dialysis_v_decile_viz # view

# Save to PDF
ggsave("/Users/edgaraguila/Desktop/School/Research Projects/UChicago Research/Data/dialysis_time_association Graphs/Median Dialysis Time v. Decile.pdf", plot = dialysis_v_decile_viz, height = 4, units = "in")
```

# % Preemptive Listing v. Decile Graph
```{r}
# Create data frame
## SVM
svm_preemp <- study_cohort_df %>%
  group_by(SVM_decile, preemptive_listing) %>% # Group by decile and result
  summarize(count = n(), .groups = "drop") %>% # Create count column
  as.data.frame() %>% # Convert to data frame
  group_by(SVM_decile) %>%
  mutate(SVM = (count / sum(count)) * 100) %>% # Create percentage column
  ungroup() %>%
  filter(preemptive_listing == 1) %>% # Keep only preemptive listing
  select(SVM_decile, SVM) # Keep columns of interest
svm_preemp

## SVI
svi_preemp <- study_cohort_df %>%
  group_by(SVI_decile, preemptive_listing) %>% # Group by decile and result
  summarize(count = n(), .groups = "drop") %>% # Create count column
  as.data.frame() %>% # Convert to data frame
  group_by(SVI_decile) %>%
  mutate(SVI = (count / sum(count)) * 100) %>% # Create percentage column
  ungroup() %>%
  filter(preemptive_listing == 1) %>% # Keep only preemptive listing
  select(SVI_decile, SVI)  # Keep columns of interest
svi_preemp

## ADI
adi_preemp <- study_cohort_df %>%
  group_by(ADI_decile, preemptive_listing) %>% # Group by decile and result
  summarize(count = n(), .groups = "drop") %>% # Create count column
  as.data.frame() %>% # Convert to data frame
  group_by(ADI_decile) %>%
  mutate(ADI = (count / sum(count)) * 100) %>% # Create percentage column
  ungroup() %>%
  filter(preemptive_listing == 1) %>% # Keep only preemptive listing
  select(ADI_decile, ADI)  # Keep columns of interest
adi_preemp

## DCI
dci_preemp <- study_cohort_df %>%
  group_by(DCI_decile, preemptive_listing) %>% # Group by decile and result
  summarize(count = n(), .groups = "drop") %>%# Create count column
  as.data.frame() %>% # Convert to data frame
  group_by(DCI_decile) %>%
  mutate(DCI = (count / sum(count)) * 100) %>% # Create percentage column
  ungroup() %>%
  filter(preemptive_listing == 1) %>% # Keep only preemptive listing
  select(DCI_decile, DCI)  # Keep columns of interest
dci_preemp

# Combine all data sets
preemp_df <- cbind(svm_preemp, svi_preemp, adi_preemp, dci_preemp) %>% # Bind by columns
  select(-SVI_decile, -ADI_decile, -DCI_decile) %>% # Remove extra decile columns
  rename(decile = SVM_decile) # Rename to have single decile column
preemp_df

rm(svm_preemp, svi_preemp, adi_preemp, dci_preemp) # Remove intermediate data frames

# Reshape data
preemp_df <- preemp_df %>% 
  pivot_longer(cols = c(SVM, SVI, ADI, DCI), # Turn the data for easier graphing 
               names_to = "index", # Set name of columns to "index"
               values_to = "percent") # Set value of columns to "percent"

# Create bar graph w/o facets
preemp_df_visual <- ggplot(preemp_df, aes(x = decile, y = percent, fill = index)) + # Set base
  geom_col(position = "dodge") + # Columns to be side-by-side
  labs(title = "Percentage of Preemptive Listing by Decile", # Set title
       subtitle = "Data from SRTR 2024", # Set subtitle
       x = "Decile", # Set x-axis label
       y = "Percentage", # Set y-axis label
       fill = "Index") + # Set legend title
  theme_classic() + # Set theme (preference)
  scale_y_continuous(limits = c(0, 60), breaks = seq(0, 60, by = 10))
preemp_df_visual # view

# Create bar graph w/ facets
preemp_df_visual_facet <- ggplot(preemp_df, aes(x = decile, y = percent, fill = index)) + # Set base
  geom_col(position = "dodge") + # Columns to be side-by-side
  facet_wrap(~ index) + # Create 4 facet grid
  labs(title = "Percentage of Preemptive Listing by Decile", # Set title
       subtitle = "Data from SRTR 2024", # Set subtitle
       x = "Decile", # Set x-axis label
       y = "Percentage", # Set y-axis label
       fill = "Index") + # Set legend title
  theme_bw() + # Set theme (preference)
  scale_y_continuous(limits = c(0, 60), breaks = seq(0, 60, by = 10))
preemp_df_visual_facet # view

# Save to PDF
ggsave("/Users/edgaraguila/Desktop/School/Research Projects/UChicago Research/Data/dialysis_time_association Graphs/Preemptive Listing Percentages.pdf", plot = preemp_df_visual, height = 4, units = "in")

ggsave("/Users/edgaraguila/Desktop/School/Research Projects/UChicago Research/Data/dialysis_time_association Graphs/Preemptive Listing Percentages facets.pdf", plot = preemp_df_visual_facet, height = 4, units = "in")
```

# Mean EPTS Score v. Decile Boxplot (Working)
```{r}
# SVM
epts_svm <- study_cohort_df %>%
  group_by(SVM_decile) %>% # Group by decile
  summarize(
    svm_lower = min(raw_epts), # 
    svm_ymin = quantile(raw_epts, 0.25),
    svm_mean = mean(raw_epts),
    svm_ymax = quantile(raw_epts, 0.75),
    svm_upper = max(raw_epts)) %>% 
  as.data.frame() # Convert to data frame

# SVI
epts_svi <- study_cohort_df %>%
  group_by(SVI_decile) %>% # Group by decile
  summarize(
    svi_lower = min(raw_epts),
    svi_ymin = quantile(raw_epts, 0.25),
    svi_mean = mean(raw_epts),
    svi_ymax = quantile(raw_epts, 0.75),
    svi_upper = max(raw_epts)) %>% 
  as.data.frame() # Convert to data frame

# ADI
epts_adi <- study_cohort_df %>%
  group_by(ADI_decile) %>% # Group by decile
  summarize(
    adi_lower = min(raw_epts),
    adi_ymin = quantile(raw_epts, 0.25),
    adi_mean = mean(raw_epts),
    adi_ymax = quantile(raw_epts, 0.75),
    adi_upper = max(raw_epts)) %>% 
  as.data.frame() # Convert to data frame

# DCI
epts_dci <- study_cohort_df %>%
  group_by(DCI_decile) %>% # Group by decile
  summarize(
    dci_lower = min(raw_epts),
    dci_ymin = quantile(raw_epts, 0.25),
    dci_mean = mean(raw_epts),
    dci_ymax = quantile(raw_epts, 0.75),
    dci_upper = max(raw_epts)) %>% 
  as.data.frame() # Convert to data frame

# Combine data frames
epts_df <- cbind(epts_svm, epts_svi, epts_adi, epts_dci) %>% # Bind by columns
  select(-SVI_decile, -ADI_decile, -DCI_decile) %>% # Remove extra decile columns
  rename(decile = SVM_decile) # Rename to have single decile column
epts_df

rm(epts_adi, epts_dci, epts_svi, epts_svm) # Remove excess data frames

epts_df <- epts_df %>%
  pivot_longer(cols = starts_with("svm") | starts_with("svi") | starts_with("adi") | starts_with("dci"),
               names_to = c("index", "stat"),
               names_pattern = "(.*)_(.*)",
               values_to = "value") %>%
  pivot_wider(names_from = stat,
              values_from = value)

epts_df$decile <- as.factor(epts_df$decile) # Convert to factor

facet_labels <- c(
  adi = "ADI",
  dci = "DCI",
  svi = "SVI",
  svm = "SVM"
)

# Graph Mean EPTS v. Decile
epts_df_visual <- ggplot(epts_df, aes(x = decile, fill = index)) + # Set base
  geom_boxplot(aes(
    ymin = lower,
    lower = ymin,
    middle = mean,
    upper = ymax,
    ymax = upper
  ), stat = "identity") +
  facet_wrap(~ index, labeller = labeller(index = facet_labels)) + # Create facet grid
  theme_bw() + # Set theme 
  theme(legend.position = "none") + # Remove legend
  labs(title = "Mean EPTS Score by Decile",
       subtitle = "Data from SRTR 2024 and EPTS Database",
       x = "Decile",
       y = "EPTS Score") 
epts_df_visual # view

# Save to PDF
ggsave("/Users/edgaraguila/Desktop/School/Research Projects/UChicago Research/Data/dialysis_time_association Graphs/Mean EPTS v. Decile.pdf", plot = epts_df_visual, height = 5, units = "in")
```

# Top 20% EPTS v. Decile
```{r}
# Create data frame
## SVM
epts_20_svm <- study_cohort_df %>%
  group_by(SVM_decile, top_percentile_epts) %>% # Group by decile and result
  summarize(count = n(), .groups = "drop") %>% # Create count column
  as.data.frame() %>% # Convert to data frame
  group_by(SVM_decile) %>%
  mutate(SVM = (count / sum(count)) * 100) %>% # Create percentage column
  ungroup() %>%
  filter(top_percentile_epts == 1) %>% # Keep only preemptive listing
  select(SVM_decile, SVM) # Keep columns of interest
epts_20_svm

## SVI
epts_20_svi <- study_cohort_df %>%
  group_by(SVI_decile, top_percentile_epts) %>% # Group by decile and result
  summarize(count = n(), .groups = "drop") %>% # Create count column
  as.data.frame() %>% # Convert to data frame
  group_by(SVI_decile) %>%
  mutate(SVI = (count / sum(count)) * 100) %>% # Create percentage column
  ungroup() %>%
  filter(top_percentile_epts == 1) %>% # Keep only preemptive listing
  select(SVI_decile, SVI)  # Keep columns of interest
epts_20_svi

## ADI
epts_20_adi <- study_cohort_df %>%
  group_by(ADI_decile, top_percentile_epts) %>% # Group by decile and result
  summarize(count = n(), .groups = "drop") %>% # Create count column
  as.data.frame() %>% # Convert to data frame
  group_by(ADI_decile) %>%
  mutate(ADI = (count / sum(count)) * 100) %>% # Create percentage column
  ungroup() %>%
  filter(top_percentile_epts == 1) %>% # Keep only preemptive listing
  select(ADI_decile, ADI)  # Keep columns of interest
epts_20_adi

## DCI
epts_20_dci <- study_cohort_df %>%
  group_by(DCI_decile, top_percentile_epts) %>% # Group by decile and result
  summarize(count = n(), .groups = "drop") %>%# Create count column
  as.data.frame() %>% # Convert to data frame
  group_by(DCI_decile) %>%
  mutate(DCI = (count / sum(count)) * 100) %>% # Create percentage column
  ungroup() %>%
  filter(top_percentile_epts == 1) %>% # Keep only preemptive listing
  select(DCI_decile, DCI)  # Keep columns of interest
epts_20_dci

# Combine all data sets
epts_20_df <- cbind(epts_20_svm, epts_20_svi, epts_20_adi, epts_20_dci) %>% # Bind by columns
  select(-SVI_decile, -ADI_decile, -DCI_decile) %>% # Remove extra decile columns
  rename(decile = SVM_decile) # Rename to have single decile column
epts_20_df

rm(epts_20_svm, epts_20_svi, epts_20_adi, epts_20_dci)

# Reshape data
epts_20_df <- epts_20_df %>% 
  pivot_longer(cols = c(SVM, SVI, ADI, DCI), # Turn the data for easier graphing (columns to rows)
               names_to = "index", # Set name of columns to "index"
               values_to = "percent") # Set value of columns to "percent"

# Create bar graph
epts_20_df_visual <- ggplot(epts_20_df, aes(x = decile, y = percent, fill = index)) + # Set base
  geom_col(position = "dodge") + # Columns to be side-by-side
  labs(title = "Percentage of Top 20% EPTS Score by Decile", # Set title
       subtitle = "Data from SRTR 2024 and EPTS Database", # Set subtitle
       x = "Decile", # Set x-axis label
       y = "Percentage", # Set y-axis label
       fill = "Index") + # Set legend title
  theme_classic() + # Set theme (preference)
  scale_y_continuous(limits = c(0, 40), breaks = seq(0, 40, by = 10))
epts_20_df_visual # view

# Save to PDF
ggsave("/Users/edgaraguila/Desktop/School/Research Projects/UChicago Research/Data/dialysis_time_association Graphs/Top 20 EPTS Graph.pdf", plot = epts_20_df_visual, height = 5, units = "in")
```

# Zero-Inflated Negative Binomial Model
```{r}
#install.packages("pscl") # Unhashtag to activate
library(pscl)

# Convert dialysis_time_at_list back to days
study_cohort_df <- study_cohort_df %>%
  mutate(dialysis_time_at_list = case_when(
    is.na(CAN_DIAL_DT) | as.Date(min_list_date) < as.Date(CAN_DIAL_DT) ~ 0, # Input 0 for patients never on dialysis
    T ~ as.numeric(difftime(min_list_date, CAN_DIAL_DT))
  ))

# Reset baseline to 5th decile
study_cohort_df$SVM_decile <- relevel(study_cohort_df$SVM_decile, ref = "5")
study_cohort_df$SVI_decile <- relevel(study_cohort_df$SVI_decile, ref = "5")
study_cohort_df$ADI_decile <- relevel(study_cohort_df$ADI_decile, ref = "5")
study_cohort_df$DCI_decile <- relevel(study_cohort_df$DCI_decile, ref = "5")

# Create model
## SVM
ZINB_svm <- zeroinfl(dialysis_time_at_list ~ SVM_decile | SVM_decile, data = study_cohort_df, dist = "negbin")
ZINB_svm

## SVI
ZINB_svi <- zeroinfl(dialysis_time_at_list ~ SVI_decile | SVI_decile, data = study_cohort_df, , dist = "negbin")
ZINB_svi

## SVM
ZINB_adi <- zeroinfl(dialysis_time_at_list ~ ADI_decile | ADI_decile, data = study_cohort_df, dist = "negbin")
ZINB_adi

## SVM
ZINB_dci <- zeroinfl(dialysis_time_at_list ~ DCI_decile | DCI_decile, data = study_cohort_df, dist = "negbin")
ZINB_dci
```

# ZINB v. Decile plots
```{r}
# Create new data frame
decile <- c(1:4, 6:10)

zinb_df <- data.frame(decile = decile) # Convert to data frame
zinb_df # view

# Create Coefficient columns
zinb_df <- zinb_df %>%
  mutate(
    SVM_ct = exp(coef(ZINB_svm, model = "count")[-1]), # SVM count
    SVM_0 = exp(coef(ZINB_svm, model = "zero")[-1]), # SVM zero
    SVI_ct = exp(coef(ZINB_svi, model = "count")[-1]), # SVI count
    SVI_0 = exp(coef(ZINB_svi, model = "zero")[-1]), # SVI zero
    ADI_ct = exp(coef(ZINB_adi, model = "count")[-1]), # ADI count
    ADI_0 = exp(coef(ZINB_adi, model = "zero")[-1]), # ADI zero
    DCI_ct = exp(coef(ZINB_dci, model = "count")[-1]), # DCI count
    DCI_0 = exp(coef(ZINB_dci, model = "zero")[-1]), # DCI zero
  )
zinb_df # view

# Insert baseline row
zinb_df <- zinb_df %>%
  rbind(data.frame(decile = 5, SVM_ct = 1, SVM_0 = 1, SVI_ct = 1, SVI_0 = 1, ADI_ct = 1, ADI_0 = 1, DCI_ct = 1, DCI_0 = 1))

zinb_df <- zinb_df[order(zinb_df$decile), ] # Put decile column in order
zinb_df # view

# Rehspae Data Frame
zinb_df <- zinb_df %>%
  pivot_longer(cols = c(SVM_ct, SVM_0, SVI_ct, SVI_0, ADI_ct, ADI_0, DCI_ct, DCI_0),
               names_to = "index",
               values_to = "coeff") %>%
  mutate(model = case_when(
    grepl("_ct$", index) ~ "Dialysis Time at Listing", # If index value ends in ct, set value to count
    grepl("_0$", index) ~ "Preemptive Listing", # If index value ends in 0, set value to zero
  ),
        index_c = case_when(
          grepl("^ADI_", index) ~ "ADI", # If index begins with ADI, set index_c (for graphing color) to ADI
          grepl("^SVI_", index) ~ "SVI", # If index begins with SVI, set index_c (for graphing color) to SVI
          grepl("^SVM_", index) ~ "SVM", # If index begins with SVM, set index_c (for graphing color) to SVM
          grepl("^DCI_", index) ~ "DCI" # If index begins with DCI, set index_c (for graphing color) to DCI
        ))

# Create Graph
zinb_df_visual <- ggplot(zinb_df, aes(x = decile, y = coeff, color = index_c)) + # Set base
  geom_line() + # Set graphing
  theme_bw() + # Set theme (preference)
  facet_wrap(~ model, scales = "free_y", ncol = 1) + # Create grid with different y-axis and stacked
  labs(title = "Zero-Inflated Negative Binomial Model of Dialysis Time and Preemptive Listing", # Set title
       subtitle = "Data from SRTR 2024", # Set subtitle
       x = "Decile", # Set x-axis label
       y = "Exponentiated Correlation Coefficient", # Set y-axis label
       color = "Index") + # Set legend title
  scale_x_continuous(limits = c(1, 10), breaks = seq(1, 10, by = 1)) # Create x-axis to show all 10 numbers
zinb_df_visual # view

# Save to PDF
ggsave("/Users/edgaraguila/Desktop/School/Research Projects/UChicago Research/Data/dialysis_time_association Graphs/ZINB Exponentiated Models.pdf", plot = zinb_df_visual, height = 4, units = "in")
```