---
title: "Regressions"
date: "2024-08-03"
output: html_document
---

##code from epts chunks -14,514 cases where death == 0 but death_outcome == 1. these should be be those removed for death/deterioration
```{r}
df_cand_kipa_w_scores %>% filter(death == 0) %>% filter(death_outcome == 1)
```
```{r}
df_cand_kipa_w_scores %>% filter(death_outcome == 1) %>% filter(transplant_outcome == 1) ##no overlap - this is good
```

##Cox model - SVM transplant
```{r}
library(survival)
df_cand_kipa_w_scores$survival_time <- as.numeric(df_cand_kipa_w_scores$survival_time)
df_cand_kipa_w_scores$time <- df_cand_kipa_w_scores$survival_time
# Convert transplant_outcome to numeric (if needed)
df_cand_kipa_w_scores$transplant_outcome <- as.numeric(df_cand_kipa_w_scores$transplant_outcome)

# Directly copy transplant_outcome to status
df_cand_kipa_w_scores$status <- df_cand_kipa_w_scores$transplant_outcome

# Fit the unadjusted Cox model
unadjusted_model <- coxph(Surv(time, status) ~ SVM_decile, data = df_cand_kipa_w_scores)

# Fit the adjusted Cox model
adjusted_model <- coxph(Surv(time, status) ~ SVM_decile + CAN_GENDER + CAN_AGE_AT_LISTING + dial_time + percentile_epts, data = df_cand_kipa_w_scores)

# Load necessary packages
library(broom)

# Extract results using broom
unadjusted_results <- tidy(unadjusted_model) %>%
  select(term, estimate, std.error, p.value) %>%
  mutate(model = "Unadjusted")

adjusted_results <- tidy(adjusted_model) %>%
  select(term, estimate, std.error, p.value) %>%
  mutate(model = "Adjusted")

# Combine results into one data frame
combined_results <- bind_rows(unadjusted_results, adjusted_results)

# Calculate Hazard Ratios (HR) and Confidence Intervals (CI)
combined_results <- combined_results %>%
  mutate(
    HR = exp(estimate),
    lower_CI = exp(estimate - 1.96 * std.error),
    upper_CI = exp(estimate + 1.96 * std.error)
  )

# Print results
library(dplyr)

combined_results %>%
  arrange(term, model) %>%
  select(term, model, HR, lower_CI, upper_CI, p.value) %>%
  rename(
    Variable = term,
    Model = model,
    "Hazard Ratio (HR)" = HR,
    "95% CI Lower" = lower_CI,
    "95% CI Upper" = upper_CI,
    "P-value" = p.value
  )

# Export to CSV
write.csv(combined_results, file = "cox_model_results_svm.csv", row.names = FALSE)
```
##Cox model - SVM death/deterioration
```{r}
df_cand_kipa_w_scores$survival_time <- as.numeric(df_cand_kipa_w_scores$survival_time)
df_cand_kipa_w_scores$time <- df_cand_kipa_w_scores$survival_time
# Convert transplant_outcome to numeric (if needed)
df_cand_kipa_w_scores$death_outcome <- as.numeric(df_cand_kipa_w_scores$death_outcome)

# Directly copy transplant_outcome to status
df_cand_kipa_w_scores$status <- df_cand_kipa_w_scores$death_outcome

# Fit the unadjusted Cox model
unadjusted_model <- coxph(Surv(time, status) ~ SVM_decile, data = df_cand_kipa_w_scores)

# Fit the adjusted Cox model
adjusted_model <- coxph(Surv(time, status) ~ SVM_decile + CAN_GENDER + CAN_AGE_AT_LISTING + dial_time + percentile_epts, data = df_cand_kipa_w_scores)

# Load necessary packages
library(broom)

# Extract results using broom
unadjusted_results <- tidy(unadjusted_model) %>%
  select(term, estimate, std.error, p.value) %>%
  mutate(model = "Unadjusted")

adjusted_results <- tidy(adjusted_model) %>%
  select(term, estimate, std.error, p.value) %>%
  mutate(model = "Adjusted")

# Combine results into one data frame
combined_results <- bind_rows(unadjusted_results, adjusted_results)

# Calculate Hazard Ratios (HR) and Confidence Intervals (CI)
combined_results <- combined_results %>%
  mutate(
    HR = exp(estimate),
    lower_CI = exp(estimate - 1.96 * std.error),
    upper_CI = exp(estimate + 1.96 * std.error)
  )

# Print results
library(dplyr)

combined_results %>%
  arrange(term, model) %>%
  select(term, model, HR, lower_CI, upper_CI, p.value) %>%
  rename(
    Variable = term,
    Model = model,
    "Hazard Ratio (HR)" = HR,
    "95% CI Lower" = lower_CI,
    "95% CI Upper" = upper_CI,
    "P-value" = p.value
  )

# Export to CSV
write.csv(combined_results, file = "cox_model_results_svm2.csv", row.names = FALSE)
```
##Cox model - SVI transplant
```{r}
df_cand_kipa_w_scores$SVI_decile <- as.factor(df_cand_kipa_w_scores$SVI_decile)
df_cand_kipa_w_scores$SVI_decile <- relevel(df_cand_kipa_w_scores$SVI_decile, ref = "5")

df_cand_kipa_w_scores$survival_time <- as.numeric(df_cand_kipa_w_scores$survival_time)
df_cand_kipa_w_scores$time <- df_cand_kipa_w_scores$survival_time
# Convert transplant_outcome to numeric (if needed)
df_cand_kipa_w_scores$transplant_outcome <- as.numeric(df_cand_kipa_w_scores$transplant_outcome)

# Directly copy transplant_outcome to status
df_cand_kipa_w_scores$status <- df_cand_kipa_w_scores$transplant_outcome

# Fit the unadjusted Cox model
unadjusted_model <- coxph(Surv(time, status) ~ SVI_decile, data = df_cand_kipa_w_scores)

# Fit the adjusted Cox model
adjusted_model <- coxph(Surv(time, status) ~ SVI_decile + CAN_GENDER + CAN_AGE_AT_LISTING + dial_time + percentile_epts, data = df_cand_kipa_w_scores)

# Load necessary packages
library(broom)

# Extract results using broom
unadjusted_results <- tidy(unadjusted_model) %>%
  select(term, estimate, std.error, p.value) %>%
  mutate(model = "Unadjusted")

adjusted_results <- tidy(adjusted_model) %>%
  select(term, estimate, std.error, p.value) %>%
  mutate(model = "Adjusted")

# Combine results into one data frame
combined_results <- bind_rows(unadjusted_results, adjusted_results)

# Calculate Hazard Ratios (HR) and Confidence Intervals (CI)
combined_results <- combined_results %>%
  mutate(
    HR = exp(estimate),
    lower_CI = exp(estimate - 1.96 * std.error),
    upper_CI = exp(estimate + 1.96 * std.error)
  )

# Print results
library(dplyr)

combined_results %>%
  arrange(term, model) %>%
  select(term, model, HR, lower_CI, upper_CI, p.value) %>%
  rename(
    Variable = term,
    Model = model,
    "Hazard Ratio (HR)" = HR,
    "95% CI Lower" = lower_CI,
    "95% CI Upper" = upper_CI,
    "P-value" = p.value
  )

# Export to CSV
write.csv(combined_results, file = "cox_model_results_svi.csv", row.names = FALSE)
```
##SVI kaplan meier (probably not useful)
```{r}
# Load necessary libraries
library(survival)
library(survminer)  # For better survival plots
library(RColorBrewer)

# Ensure SVI_decile is a factor and set reference level
df_cand_kipa_w_scores$SVI_decile <- as.factor(df_cand_kipa_w_scores$SVI_decile)
df_cand_kipa_w_scores$SVI_decile <- relevel(df_cand_kipa_w_scores$SVI_decile, ref = "5")

# Convert time and status variables to appropriate types
df_cand_kipa_w_scores$time <- as.numeric(df_cand_kipa_w_scores$survival_time)
df_cand_kipa_w_scores$status <- as.numeric(df_cand_kipa_w_scores$transplant_outcome)

# Fit the Kaplan-Meier model
km_fit <- survfit(Surv(time, status) ~ SVI_decile, data = df_cand_kipa_w_scores)

palette <- brewer.pal(n = length(levels(df_cand_kipa_w_scores$SVI_decile)), name = "Dark2")

# Example: Zooming in on a specific time range and survival probability range

# Define your zoom-in ranges
xlim_range <- c(0, 1850)  # Adjust the time range as needed
ylim_range <- c(0, 1)    # Adjust the survival probability range as needed

# Plot Kaplan-Meier curves with adjusted zoom-in ranges
ggsurvplot(km_fit, data = df_cand_kipa_w_scores,
           pval = TRUE,               # Show p-value of the log-rank test
           conf.int = TRUE,           # Show confidence intervals
           risk.table = TRUE,         # Show the number of at risk individuals
           title = "Kaplan-Meier Survival Curves by SVI Decile",
           xlab = "Time",
           ylab = "Survival Probability",
           xlim = xlim_range,         # Set x-axis limits
           ylim = ylim_range)         # Set y-axis limits

```

```{r}
# Load necessary packages
library(survminer)
library(survival)

# Fit Kaplan-Meier model
fit <- survfit(Surv(time, status) ~ SVI_decile, data = df_cand_kipa_w_scores)

# Plot Kaplan-Meier survival curves
ggsurvplot(
  fit,
  data = df_cand_kipa_w_scores,
  size = 0.5,
  pval = TRUE,
  legend.labs = levels(df_cand_kipa_w_scores$SVI_decile),
  palette = "Dark2",
  ggtheme = theme_minimal()
)
```
##HR scatter plot
```{r}
# Create a scatter plot with HR
ggplot(combined_results, aes(x = term, y = HR, color = model)) +
  geom_point(aes(size = HR)) +
  geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), width = 0.2) +
  scale_y_log10() +  # Use log scale for better visualization of ratios
  labs(
    x = "Variable",
    y = "Hazard Ratio (HR)",
    title = "Scatter Plot of Hazard Ratios"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# Fit the Cox proportional hazards model
cox_fit <- coxph(Surv(time, status) ~ SVI_decile + CAN_GENDER + CAN_AGE_AT_LISTING + dial_time + percentile_epts, data = df_cand_kipa_w_scores)

# Predict survival curves for different levels of SVI_decile
# Create a new data frame with specific values
new_data <- data.frame(
  SVI_decile = factor(c("1", "2", "3", "4", "6", "7", "8", "9", "10"), levels = levels(df_cand_kipa_w_scores$SVI_decile)),
  CAN_GENDER = df_cand_kipa_w_scores$CAN_GENDER,
  CAN_AGE_AT_LISTING = mean(df_cand_kipa_w_scores$CAN_AGE_AT_LISTING, na.rm = TRUE),
  dial_time = mean(df_cand_kipa_w_scores$dial_time, na.rm = TRUE),
  percentile_epts = mean(df_cand_kipa_w_scores$percentile_epts, na.rm = TRUE)
)

# Predict survival curves
surv_fit_adjusted <- survfit(cox_fit, newdata = new_data)

# Plot adjusted survival curves
ggsurvplot(surv_fit_adjusted, data = new_data,
           pval = TRUE,
           conf.int = TRUE,
           title = "Adjusted Survival Curves by SVI Decile",
           xlab = "Time",
           ylab = "Survival Probability")
```
##HR forest plot
```{r}
# Load necessary packages
library(ggplot2)
library(dplyr)

# Create the forest plot
ggplot(combined_results, aes(x = HR, ymin = lower_CI, ymax = upper_CI, color = model)) +
  geom_pointrange(aes(y = term)) +
  scale_x_log10() +  # Use log scale for better visualization of ratios
  labs(
    x = "Hazard Ratio (HR)",
    y = "Variable",
    title = "Forest Plot of Hazard Ratios"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.y = element_text(size = 10)
  )
```
```{r}
df_cand_kipa_w_scores %>% filter(status == "censor")

df_cand_kipa_w_scores %>% filter(status == "transplant")

df_cand_kipa_w_scores %>% filter(status == "death/deteriorate")
```

##Cox model - SVI death/deterioration
```{r}
df_cand_kipa_w_scores$SVI_decile <- as.factor(df_cand_kipa_w_scores$SVI_decile)
df_cand_kipa_w_scores$SVI_decile <- relevel(df_cand_kipa_w_scores$SVI_decile, ref = "5")

df_cand_kipa_w_scores$survival_time <- as.numeric(df_cand_kipa_w_scores$survival_time)
df_cand_kipa_w_scores$time <- df_cand_kipa_w_scores$survival_time
# Convert transplant_outcome to numeric (if needed)
df_cand_kipa_w_scores$death_outcome <- as.numeric(df_cand_kipa_w_scores$death_outcome)

# Directly copy transplant_outcome to status
df_cand_kipa_w_scores$status <- df_cand_kipa_w_scores$death_outcome

# Fit the unadjusted Cox model
unadjusted_model <- coxph(Surv(time, status) ~ SVI_decile, data = df_cand_kipa_w_scores)

# Fit the adjusted Cox model
adjusted_model <- coxph(Surv(time, status) ~ SVI_decile + CAN_GENDER + CAN_AGE_AT_LISTING + dial_time + percentile_epts, data = df_cand_kipa_w_scores)

# Load necessary packages
library(broom)

# Extract results using broom
unadjusted_results <- tidy(unadjusted_model) %>%
  select(term, estimate, std.error, p.value) %>%
  mutate(model = "Unadjusted")

adjusted_results <- tidy(adjusted_model) %>%
  select(term, estimate, std.error, p.value) %>%
  mutate(model = "Adjusted")

# Combine results into one data frame
combined_results <- bind_rows(unadjusted_results, adjusted_results)

# Calculate Hazard Ratios (HR) and Confidence Intervals (CI)
combined_results <- combined_results %>%
  mutate(
    HR = exp(estimate),
    lower_CI = exp(estimate - 1.96 * std.error),
    upper_CI = exp(estimate + 1.96 * std.error)
  )

# Print results
library(dplyr)

combined_results %>%
  arrange(term, model) %>%
  select(term, model, HR, lower_CI, upper_CI, p.value) %>%
  rename(
    Variable = term,
    Model = model,
    "Hazard Ratio (HR)" = HR,
    "95% CI Lower" = lower_CI,
    "95% CI Upper" = upper_CI,
    "P-value" = p.value
  )

# Export to CSV
write.csv(combined_results, file = "cox_model_results_svi2.csv", row.names = FALSE)
```
##Cox model - ADI transplant
```{r}
df_cand_kipa_w_scores$survival_time <- as.numeric(df_cand_kipa_w_scores$survival_time)
df_cand_kipa_w_scores$time <- df_cand_kipa_w_scores$survival_time
# Convert transplant_outcome to numeric (if needed)
df_cand_kipa_w_scores$transplant_outcome <- as.numeric(df_cand_kipa_w_scores$transplant_outcome)

# Directly copy transplant_outcome to status
df_cand_kipa_w_scores$status <- df_cand_kipa_w_scores$transplant_outcome

# Fit the unadjusted Cox model
unadjusted_model <- coxph(Surv(time, status) ~ ADI_decile, data = df_cand_kipa_w_scores)

# Fit the adjusted Cox model
adjusted_model <- coxph(Surv(time, status) ~ ADI_decile + CAN_GENDER + CAN_AGE_AT_LISTING + dial_time + percentile_epts, data = df_cand_kipa_w_scores)

# Load necessary packages
library(broom)

# Extract results using broom
unadjusted_results <- tidy(unadjusted_model) %>%
  select(term, estimate, std.error, p.value) %>%
  mutate(model = "Unadjusted")

adjusted_results <- tidy(adjusted_model) %>%
  select(term, estimate, std.error, p.value) %>%
  mutate(model = "Adjusted")

# Combine results into one data frame
combined_results <- bind_rows(unadjusted_results, adjusted_results)

# Calculate Hazard Ratios (HR) and Confidence Intervals (CI)
combined_results <- combined_results %>%
  mutate(
    HR = exp(estimate),
    lower_CI = exp(estimate - 1.96 * std.error),
    upper_CI = exp(estimate + 1.96 * std.error)
  )

# Print results
library(dplyr)

combined_results %>%
  arrange(term, model) %>%
  select(term, model, HR, lower_CI, upper_CI, p.value) %>%
  rename(
    Variable = term,
    Model = model,
    "Hazard Ratio (HR)" = HR,
    "95% CI Lower" = lower_CI,
    "95% CI Upper" = upper_CI,
    "P-value" = p.value
  )

# Export to CSV
write.csv(combined_results, file = "cox_model_results_adi.csv", row.names = FALSE)
```
##Cox model - ADI death/deterioration
```{r}
df_cand_kipa_w_scores$survival_time <- as.numeric(df_cand_kipa_w_scores$survival_time)
df_cand_kipa_w_scores$time <- df_cand_kipa_w_scores$survival_time
# Convert transplant_outcome to numeric (if needed)
df_cand_kipa_w_scores$death_outcome <- as.numeric(df_cand_kipa_w_scores$death_outcome)

# Directly copy transplant_outcome to status
df_cand_kipa_w_scores$status <- df_cand_kipa_w_scores$death_outcome

# Fit the unadjusted Cox model
unadjusted_model <- coxph(Surv(time, status) ~ ADI_decile, data = df_cand_kipa_w_scores)

# Fit the adjusted Cox model
adjusted_model <- coxph(Surv(time, status) ~ ADI_decile + CAN_GENDER + CAN_AGE_AT_LISTING + dial_time + percentile_epts, data = df_cand_kipa_w_scores)

# Load necessary packages
library(broom)

# Extract results using broom
unadjusted_results <- tidy(unadjusted_model) %>%
  select(term, estimate, std.error, p.value) %>%
  mutate(model = "Unadjusted")

adjusted_results <- tidy(adjusted_model) %>%
  select(term, estimate, std.error, p.value) %>%
  mutate(model = "Adjusted")

# Combine results into one data frame
combined_results <- bind_rows(unadjusted_results, adjusted_results)

# Calculate Hazard Ratios (HR) and Confidence Intervals (CI)
combined_results <- combined_results %>%
  mutate(
    HR = exp(estimate),
    lower_CI = exp(estimate - 1.96 * std.error),
    upper_CI = exp(estimate + 1.96 * std.error)
  )

# Print results
library(dplyr)

combined_results %>%
  arrange(term, model) %>%
  select(term, model, HR, lower_CI, upper_CI, p.value) %>%
  rename(
    Variable = term,
    Model = model,
    "Hazard Ratio (HR)" = HR,
    "95% CI Lower" = lower_CI,
    "95% CI Upper" = upper_CI,
    "P-value" = p.value
  )

# Export to CSV
write.csv(combined_results, file = "cox_model_results_adi2.csv", row.names = FALSE)
```
##Cox model - DCI transplant
```{r}
df_cand_kipa_w_scores$survival_time <- as.numeric(df_cand_kipa_w_scores$survival_time)
df_cand_kipa_w_scores$time <- df_cand_kipa_w_scores$survival_time
# Convert transplant_outcome to numeric (if needed)
df_cand_kipa_w_scores$transplant_outcome <- as.numeric(df_cand_kipa_w_scores$transplant_outcome)

# Directly copy transplant_outcome to status
df_cand_kipa_w_scores$status <- df_cand_kipa_w_scores$transplant_outcome

# Fit the unadjusted Cox model
unadjusted_model <- coxph(Surv(time, status) ~ DCI_decile, data = df_cand_kipa_w_scores)

# Fit the adjusted Cox model
adjusted_model <- coxph(Surv(time, status) ~ DCI_decile + CAN_GENDER + CAN_AGE_AT_LISTING + dial_time + percentile_epts, data = df_cand_kipa_w_scores)

# Load necessary packages
library(broom)

# Extract results using broom
unadjusted_results <- tidy(unadjusted_model) %>%
  select(term, estimate, std.error, p.value) %>%
  mutate(model = "Unadjusted")

adjusted_results <- tidy(adjusted_model) %>%
  select(term, estimate, std.error, p.value) %>%
  mutate(model = "Adjusted")

# Combine results into one data frame
combined_results <- bind_rows(unadjusted_results, adjusted_results)

# Calculate Hazard Ratios (HR) and Confidence Intervals (CI)
combined_results <- combined_results %>%
  mutate(
    HR = exp(estimate),
    lower_CI = exp(estimate - 1.96 * std.error),
    upper_CI = exp(estimate + 1.96 * std.error)
  )

# Print results
library(dplyr)

combined_results %>%
  arrange(term, model) %>%
  select(term, model, HR, lower_CI, upper_CI, p.value) %>%
  rename(
    Variable = term,
    Model = model,
    "Hazard Ratio (HR)" = HR,
    "95% CI Lower" = lower_CI,
    "95% CI Upper" = upper_CI,
    "P-value" = p.value
  )

# Export to CSV
write.csv(combined_results, file = "cox_model_results_dci.csv", row.names = FALSE)
```
##DCI - death/deterioration
```{r}
df_cand_kipa_w_scores$survival_time <- as.numeric(df_cand_kipa_w_scores$survival_time)
df_cand_kipa_w_scores$time <- df_cand_kipa_w_scores$survival_time
# Convert transplant_outcome to numeric (if needed)
df_cand_kipa_w_scores$death_outcome <- as.numeric(df_cand_kipa_w_scores$death_outcome)

# Directly copy transplant_outcome to status
df_cand_kipa_w_scores$status <- df_cand_kipa_w_scores$death_outcome

# Fit the unadjusted Cox model
unadjusted_model <- coxph(Surv(time, status) ~ DCI_decile, data = df_cand_kipa_w_scores)

# Fit the adjusted Cox model
adjusted_model <- coxph(Surv(time, status) ~ DCI_decile + CAN_GENDER + CAN_AGE_AT_LISTING + dial_time + percentile_epts, data = df_cand_kipa_w_scores)

# Load necessary packages
library(broom)

# Extract results using broom
unadjusted_results <- tidy(unadjusted_model) %>%
  select(term, estimate, std.error, p.value) %>%
  mutate(model = "Unadjusted")

adjusted_results <- tidy(adjusted_model) %>%
  select(term, estimate, std.error, p.value) %>%
  mutate(model = "Adjusted")

# Combine results into one data frame
combined_results <- bind_rows(unadjusted_results, adjusted_results)

# Calculate Hazard Ratios (HR) and Confidence Intervals (CI)
combined_results <- combined_results %>%
  mutate(
    HR = exp(estimate),
    lower_CI = exp(estimate - 1.96 * std.error),
    upper_CI = exp(estimate + 1.96 * std.error)
  )

# Print results
library(dplyr)

combined_results %>%
  arrange(term, model) %>%
  select(term, model, HR, lower_CI, upper_CI, p.value) %>%
  rename(
    Variable = term,
    Model = model,
    "Hazard Ratio (HR)" = HR,
    "95% CI Lower" = lower_CI,
    "95% CI Upper" = upper_CI,
    "P-value" = p.value
  )

# Export to CSV
write.csv(combined_results, file = "cox_model_results_dci2.csv", row.names = FALSE)
```

```{r}
library(survival)

my_time_1 <- df_cand_kipa_w_scores$survival_time
my_status_1 <- df_cand_kipa_w_scores$transplant_outcome
my_exposure_1 <- df_cand_kipa_w_scores$SVM_decile
my_covariate1_1 <- df_cand_kipa_w_scores$CAN_AGE_AT_LISTING

library(ggplot2)

df_surv <- data.frame(
  time = my_time_1,
  status = my_status_1,
  exposure = my_exposure_1,
  covariate1 = my_covariate1_1
)

my_time_1 <- as.numeric(as.character(df_cand_kipa_w_scores$survival_time))
my_status_1 <- as.numeric(as.character(df_cand_kipa_w_scores$transplant_outcome))
my_eposure_1 <- as.numeric(as.character(df_cand_kipa_w_scores$SVM_decile))
my_covariate1_1 <- as.numeric(as.character(df_cand_kipa_w_scores$CAN_AGE_AT_LISTING))

# Unadjusted Cox model
unadjusted_cox <- coxph(Surv(time, status) ~ exposure, data = df_surv)
summary(unadjusted_cox)

# Adjusted Cox model
adjusted_cox <- coxph(Surv(time, status) ~ exposure + covariate1, data = df_surv)
summary(adjusted_cox)
```
##ADI death/deterioration hazard ratio by decile
```{r}
# Generate new data for prediction
df_cand_kipa_w_scores$ADI_decile_num <- as.numeric(df_cand_kipa_w_scores$ADI_decile)

library(rms)

# Ensure the datadist is set up for the model
dd <- datadist(df_cand_kipa_w_scores)
options(datadist = "dd")

# Fit the Cox model with restricted cubic splines
model <- cph(Surv(time, death_outcome) ~ rcs(ADI_decile_num, 4), data = df_cand_kipa_w_scores, x = TRUE, y = TRUE)


new_data <- data.frame(ADI_decile_num = seq(min(df_cand_kipa_w_scores$ADI_decile_num, na.rm = TRUE),
                                            max(df_cand_kipa_w_scores$ADI_decile_num, na.rm = TRUE),
                                            length.out = 100))
# Predict the linear predictor
lp <- predict(model, newdata = new_data, type = "lp")

# Predict survival probabilities using the model's survival function
hr <- exp(lp)

# Create the data frame for plotting
plot_data <- data.frame(
  ADI_decile_num = new_data$ADI_decile_num,
  Hazard_Ratio = hr
)

library(ggplot2)

ggplot(plot_data, aes(x = ADI_decile_num, y = Hazard_Ratio)) +
  geom_line() +
  labs(title = "Hazard Ratio by ADI Decile",
       x = "ADI Decile Number",
       y = "Hazard Ratio for Death / Deterioration") +
  theme_minimal()
```

##set reference to ADI_decile = 5
```{r}
# Load necessary libraries
library(rms)
library(ggplot2)

# Convert decile column to numeric
df_cand_kipa_w_scores$ADI_decile_num <- as.numeric(df_cand_kipa_w_scores$ADI_decile)

# Ensure the datadist is set up for the model
dd <- datadist(df_cand_kipa_w_scores)
options(datadist = "dd")

# Fit the Cox model with restricted cubic splines
model <- cph(Surv(time, death_outcome) ~ rcs(ADI_decile_num, 4), data = df_cand_kipa_w_scores, x = TRUE, y = TRUE)

# Generate new data for prediction
new_data <- data.frame(ADI_decile_num = seq(min(df_cand_kipa_w_scores$ADI_decile_num, na.rm = TRUE),
                                            max(df_cand_kipa_w_scores$ADI_decile_num, na.rm = TRUE),
                                            length.out = 100))

# Predict the linear predictor
lp <- predict(model, newdata = new_data, type = "lp")

# Predict hazard ratios
hr <- exp(lp)

# Find the hazard ratio at the reference value (e.g., decile = 5)
reference_value <- 5
reference_hr <- hr[new_data$ADI_decile_num == reference_value]

# Adjust hazard ratios so that the reference value has an HR of 1
hr_adjusted <- hr / reference_hr

# Create the data frame for plotting
plot_data <- data.frame(
  ADI_decile_num = new_data$ADI_decile_num,
  Hazard_Ratio = hr_adjusted
)

# Plot
ggplot(plot_data, aes(x = ADI_decile_num, y = Hazard_Ratio)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Add a reference line at HR = 1
  labs(title = "Hazard Ratio by ADI Decile",
       x = "ADI Decile Number",
       y = "Hazard Ratio for Death/Deterioration (Reference = 5)") +
  theme_minimal()

```
##ADI transplant hazard ratio by decile
```{r}
# Load necessary libraries
library(rms)
library(ggplot2)

# Convert decile column to numeric
df_cand_kipa_w_scores$ADI_decile_num <- as.numeric(df_cand_kipa_w_scores$ADI_decile)

# Ensure the datadist is set up for the model
dd <- datadist(df_cand_kipa_w_scores)
options(datadist = "dd")

# Fit the Cox model with restricted cubic splines
model <- cph(Surv(time, transplant_outcome) ~ rcs(ADI_decile_num, 4), data = df_cand_kipa_w_scores, x = TRUE, y = TRUE)

# Generate new data for prediction
new_data <- data.frame(ADI_decile_num = seq(min(df_cand_kipa_w_scores$ADI_decile_num, na.rm = TRUE),
                                            max(df_cand_kipa_w_scores$ADI_decile_num, na.rm = TRUE),
                                            length.out = 100))

# Predict the linear predictor
lp <- predict(model, newdata = new_data, type = "lp")

# Predict hazard ratios
hr <- exp(lp)

# Find the hazard ratio at the reference value (e.g., decile = 5)
reference_value <- 5
reference_hr <- hr[new_data$ADI_decile_num == reference_value]

# Adjust hazard ratios so that the reference value has an HR of 1
hr_adjusted <- hr / reference_hr

# Create the data frame for plotting
plot_data <- data.frame(
  ADI_decile_num = new_data$ADI_decile_num,
  Hazard_Ratio = hr_adjusted
)

# Plot
ggplot(plot_data, aes(x = ADI_decile_num, y = Hazard_Ratio)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Add a reference line at HR = 1
  labs(title = "Hazard Ratio by ADI Decile",
       x = "ADI Decile Number",
       y = "Hazard Ratio for Transplant (Reference = 5)") +
  theme_minimal()

```
##SVM death/deterioration hazard ratio by decile
```{r}
# Load necessary libraries
library(rms)
library(ggplot2)

# Convert decile column to numeric
df_cand_kipa_w_scores$SVM_decile_num <- as.numeric(df_cand_kipa_w_scores$SVM_decile)

# Ensure the datadist is set up for the model
dd <- datadist(df_cand_kipa_w_scores)
options(datadist = "dd")

# Fit the Cox model with restricted cubic splines
model <- cph(Surv(time, death_outcome) ~ rcs(SVM_decile_num, 4), data = df_cand_kipa_w_scores, x = TRUE, y = TRUE)

# Generate new data for prediction
new_data <- data.frame(SVM_decile_num = seq(min(df_cand_kipa_w_scores$SVM_decile_num, na.rm = TRUE),
                                            max(df_cand_kipa_w_scores$SVM_decile_num, na.rm = TRUE),
                                            length.out = 100))

# Predict the linear predictor
lp <- predict(model, newdata = new_data, type = "lp")

# Predict hazard ratios
hr <- exp(lp)

# Find the hazard ratio at the reference value (e.g., decile = 5)
reference_value <- 5
reference_hr <- hr[new_data$SVM_decile_num == reference_value]

# Adjust hazard ratios so that the reference value has an HR of 1
hr_adjusted <- hr / reference_hr

# Create the data frame for plotting
plot_data <- data.frame(
  SVM_decile_num = new_data$SVM_decile_num,
  Hazard_Ratio = hr_adjusted
)

# Plot
ggplot(plot_data, aes(x = SVM_decile_num, y = Hazard_Ratio)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Add a reference line at HR = 1
  labs(title = "Hazard Ratio by SVM Decile",
       x = "SVM Decile Number",
       y = "Hazard Ratio of Death/Deterioration (Reference = 5)") +
  theme_minimal()
```
##SVM transplant hazard ratio by decile
```{r}
# Load necessary libraries
library(rms)
library(ggplot2)

# Convert decile column to numeric
df_cand_kipa_w_scores$SVM_decile_num <- as.numeric(df_cand_kipa_w_scores$SVM_decile)

# Ensure the datadist is set up for the model
dd <- datadist(df_cand_kipa_w_scores)
options(datadist = "dd")

# Fit the Cox model with restricted cubic splines
model <- cph(Surv(time, transplant_outcome) ~ rcs(SVM_decile_num, 4), data = df_cand_kipa_w_scores, x = TRUE, y = TRUE)

# Generate new data for prediction
new_data <- data.frame(SVM_decile_num = seq(min(df_cand_kipa_w_scores$SVM_decile_num, na.rm = TRUE),
                                            max(df_cand_kipa_w_scores$SVM_decile_num, na.rm = TRUE),
                                            length.out = 100))

# Predict the linear predictor
lp <- predict(model, newdata = new_data, type = "lp")

# Predict hazard ratios
hr <- exp(lp)

# Find the hazard ratio at the reference value (e.g., decile = 5)
reference_value <- 5
reference_hr <- hr[new_data$SVM_decile_num == reference_value]

# Adjust hazard ratios so that the reference value has an HR of 1
hr_adjusted <- hr / reference_hr

# Create the data frame for plotting
plot_data <- data.frame(
  SVM_decile_num = new_data$SVM_decile_num,
  Hazard_Ratio = hr_adjusted
)

# Plot
ggplot(plot_data, aes(x = SVM_decile_num, y = Hazard_Ratio)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Add a reference line at HR = 1
  labs(title = "Hazard Ratio by SVM Decile",
       x = "SVM Decile Number",
       y = "Hazard Ratio of Transplant (Reference = 5)") +
  theme_minimal()
```
##SVI death/deterioration hazard ratio by decile
```{r}
# Load necessary libraries
library(rms)
library(ggplot2)

# Convert decile column to numeric
df_cand_kipa_w_scores$SVI_decile_num <- as.numeric(df_cand_kipa_w_scores$SVI_decile)

# Ensure the datadist is set up for the model
dd <- datadist(df_cand_kipa_w_scores)
options(datadist = "dd")

# Fit the Cox model with restricted cubic splines
model <- cph(Surv(time, death_outcome) ~ rcs(SVI_decile_num, 4), data = df_cand_kipa_w_scores, x = TRUE, y = TRUE)

# Generate new data for prediction
new_data <- data.frame(SVI_decile_num = seq(min(df_cand_kipa_w_scores$SVI_decile_num, na.rm = TRUE),
                                            max(df_cand_kipa_w_scores$SVI_decile_num, na.rm = TRUE),
                                            length.out = 100))

# Predict the linear predictor
lp <- predict(model, newdata = new_data, type = "lp")

# Predict hazard ratios
hr <- exp(lp)

# Find the hazard ratio at the reference value (e.g., decile = 5)
reference_value <- 5
reference_hr <- hr[new_data$SVI_decile_num == reference_value]

# Adjust hazard ratios so that the reference value has an HR of 1
hr_adjusted <- hr / reference_hr

# Create the data frame for plotting
plot_data <- data.frame(
  SVI_decile_num = new_data$SVI_decile_num,
  Hazard_Ratio = hr_adjusted
)

# Plot
ggplot(plot_data, aes(x = SVI_decile_num, y = Hazard_Ratio)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Add a reference line at HR = 1
  labs(title = "Hazard Ratio by SVI Decile",
       x = "SVI Decile Number",
       y = "Hazard Ratio for Death/Deterioration (Reference = 5)") +
  theme_minimal()
```
##SVI transplant hazard ratio by decile
```{r}
# Load necessary libraries
library(rms)
library(ggplot2)

# Convert decile column to numeric
df_cand_kipa_w_scores$SVI_decile_num <- as.numeric(df_cand_kipa_w_scores$SVI_decile)

# Ensure the datadist is set up for the model
dd <- datadist(df_cand_kipa_w_scores)
options(datadist = "dd")

# Fit the Cox model with restricted cubic splines
model <- cph(Surv(time, transplant_outcome) ~ rcs(SVI_decile_num, 4), data = df_cand_kipa_w_scores, x = TRUE, y = TRUE)

# Generate new data for prediction
new_data <- data.frame(SVI_decile_num = seq(min(df_cand_kipa_w_scores$SVI_decile_num, na.rm = TRUE),
                                            max(df_cand_kipa_w_scores$SVI_decile_num, na.rm = TRUE),
                                            length.out = 100))

# Predict the linear predictor
lp <- predict(model, newdata = new_data, type = "lp")

# Predict hazard ratios
hr <- exp(lp)

# Find the hazard ratio at the reference value (e.g., decile = 5)
reference_value <- 5
reference_hr <- hr[new_data$SVI_decile_num == reference_value]

# Adjust hazard ratios so that the reference value has an HR of 1
hr_adjusted <- hr / reference_hr

# Create the data frame for plotting
plot_data <- data.frame(
  SVI_decile_num = new_data$SVI_decile_num,
  Hazard_Ratio = hr_adjusted
)

# Plot
ggplot(plot_data, aes(x = SVI_decile_num, y = Hazard_Ratio)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Add a reference line at HR = 1
  labs(title = "Hazard Ratio of Transplant by SVI Decile",
       x = "SVI Decile Number",
       y = "Hazard Ratio of Transplant (Reference = 5)") +
  theme_minimal()
```
##DCI death/deterioration hazard ratio by decile
```{r}
# Load necessary libraries
library(rms)
library(ggplot2)

# Convert decile column to numeric
df_cand_kipa_w_scores$DCI_decile_num <- as.numeric(df_cand_kipa_w_scores$DCI_decile)

# Ensure the datadist is set up for the model
dd <- datadist(df_cand_kipa_w_scores)
options(datadist = "dd")

# Fit the Cox model with restricted cubic splines
model <- cph(Surv(time, death_outcome) ~ rcs(DCI_decile_num, 4), data = df_cand_kipa_w_scores, x = TRUE, y = TRUE)

# Generate new data for prediction
new_data <- data.frame(DCI_decile_num = seq(min(df_cand_kipa_w_scores$DCI_decile_num, na.rm = TRUE),
                                            max(df_cand_kipa_w_scores$DCI_decile_num, na.rm = TRUE),
                                            length.out = 100))

# Predict the linear predictor
lp <- predict(model, newdata = new_data, type = "lp")

# Predict hazard ratios
hr <- exp(lp)

# Find the hazard ratio at the reference value (e.g., decile = 5)
reference_value <- 5
reference_hr <- hr[new_data$DCI_decile_num == reference_value]

# Adjust hazard ratios so that the reference value has an HR of 1
hr_adjusted <- hr / reference_hr

# Create the data frame for plotting
plot_data <- data.frame(
  DCI_decile_num = new_data$DCI_decile_num,
  Hazard_Ratio = hr_adjusted
)

# Plot
ggplot(plot_data, aes(x = DCI_decile_num, y = Hazard_Ratio)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Add a reference line at HR = 1
  labs(title = "Hazard Ratio by DCI Decile",
       x = "DCI Decile Number",
       y = "Hazard Ratio of Death/Deterioration (Reference = 5)") +
  theme_minimal()
```
##DCI transplant hazard ratio by decile
```{r}
# Load necessary libraries
library(rms)
library(ggplot2)

# Convert decile column to numeric
df_cand_kipa_w_scores$DCI_decile_num <- as.numeric(df_cand_kipa_w_scores$DCI_decile)

# Ensure the datadist is set up for the model
dd <- datadist(df_cand_kipa_w_scores)
options(datadist = "dd")

# Fit the Cox model with restricted cubic splines
model <- cph(Surv(time, transplant_outcome) ~ rcs(DCI_decile_num, 4), data = df_cand_kipa_w_scores, x = TRUE, y = TRUE)

# Generate new data for prediction
new_data <- data.frame(DCI_decile_num = seq(min(df_cand_kipa_w_scores$DCI_decile_num, na.rm = TRUE),
                                            max(df_cand_kipa_w_scores$DCI_decile_num, na.rm = TRUE),
                                            length.out = 100))

# Predict the linear predictor
lp <- predict(model, newdata = new_data, type = "lp")

# Predict hazard ratios
hr <- exp(lp)

# Find the hazard ratio at the reference value (e.g., decile = 5)
reference_value <- 5
reference_hr <- hr[new_data$DCI_decile_num == reference_value]

# Adjust hazard ratios so that the reference value has an HR of 1
hr_adjusted <- hr / reference_hr

# Create the data frame for plotting
plot_data <- data.frame(
  DCI_decile_num = new_data$DCI_decile_num,
  Hazard_Ratio = hr_adjusted
)

# Plot
ggplot(plot_data, aes(x = DCI_decile_num, y = Hazard_Ratio)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Add a reference line at HR = 1
  labs(title = "Hazard Ratio by DCI Decile",
       x = "DCI Decile Number",
       y = "Hazard Ratio (Reference = 5)") +
  theme_minimal()
```

```{r}
install.packages("patchwork")

```
```{r}
library(ggplot2)
library(patchwork)

```

##plot including all 4 indices - death/deterioration
```{r}
# Load necessary libraries
library(rms)
library(ggplot2)
library(dplyr)

# Define a function to generate hazard ratio data for different deciles
generate_plot_data <- function(decile_col, reference_value = 5) {
  # Convert decile column to numeric
  df_cand_kipa_w_scores[[paste0(decile_col, "_num")]] <- as.numeric(df_cand_kipa_w_scores[[decile_col]])

  # Ensure the datadist is set up for the model
  dd <- datadist(df_cand_kipa_w_scores)
  options(datadist = "dd")

  # Fit the Cox model with restricted cubic splines
  formula <- as.formula(paste("Surv(time, death_outcome) ~ rcs(", paste0(decile_col, "_num"), ", 4)"))
  model <- cph(formula, data = df_cand_kipa_w_scores, x = TRUE, y = TRUE)

  # Generate new data for prediction
  new_data <- data.frame(
    Decile = seq(min(df_cand_kipa_w_scores[[paste0(decile_col, "_num")]], na.rm = TRUE),
                 max(df_cand_kipa_w_scores[[paste0(decile_col, "_num")]], na.rm = TRUE),
                 length.out = 100)
  )
  names(new_data) <- paste0(decile_col, "_num")

  # Predict the linear predictor
  lp <- predict(model, newdata = new_data, type = "lp")

  # Predict hazard ratios
  hr <- exp(lp)

  # Find the hazard ratio at the reference value (e.g., decile = 5)
  reference_hr <- hr[new_data[[paste0(decile_col, "_num")]] == reference_value]

  # Adjust hazard ratios so that the reference value has an HR of 1
  hr_adjusted <- hr / reference_hr

  # Create the data frame for plotting
  plot_data <- data.frame(
    Decile = new_data[[paste0(decile_col, "_num")]],
    Hazard_Ratio = hr_adjusted,
    Variable = decile_col  # Add a column for the variable name
  )
  
  return(plot_data)
}

# Generate plot data for each decile
plot_data_ADI <- generate_plot_data("ADI_decile")
plot_data_SVM <- generate_plot_data("SVM_decile")
plot_data_SVI <- generate_plot_data("SVI_decile")
plot_data_DCI <- generate_plot_data("DCI_decile")

# Combine all plot data into one data frame
combined_plot_data <- bind_rows(plot_data_ADI, plot_data_SVM, plot_data_SVI, plot_data_DCI)

# Plot all lines on the same graph with x-axis intervals of 1
ggplot(combined_plot_data, aes(x = Decile, y = Hazard_Ratio, color = Variable)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Add a reference line at HR = 1
  scale_x_continuous(breaks = seq(min(combined_plot_data$Decile, na.rm = TRUE),
                                  max(combined_plot_data$Decile, na.rm = TRUE),
                                  by = 1)) +  # Set x-axis breaks to intervals of 1
  labs(title = "Hazard Ratio by Decile for Various Metrics",
       x = "Decile Number",
       y = "Hazard Ratio of Death/Deterioration(Reference = 5)",
       color = "Variable") +
  theme_minimal()

```
```{r}
# Convert decile columns to numeric
df_cand_kipa_w_scores$ADI_decile_num <- as.numeric(df_cand_kipa_w_scores$ADI_decile)
df_cand_kipa_w_scores$SVM_decile_num <- as.numeric(df_cand_kipa_w_scores$SVM_decile)
df_cand_kipa_w_scores$SVI_decile_num <- as.numeric(df_cand_kipa_w_scores$SVI_decile)
df_cand_kipa_w_scores$DCI_decile_num <- as.numeric(df_cand_kipa_w_scores$DCI_decile)
```

```{r}
# Define the formula for the model
test_formula <- as.formula("Surv(time, death_outcome) ~ rcs(ADI_decile_num, 4)")

# Fit the Cox model
test_model <- cph(test_formula, data = df_cand_kipa_w_scores, x = TRUE, y = TRUE)

# Print model summary
summary(test_model)

```
##all 4 indices - death/deterioration - with confidence intervals
```{r}
# Load necessary libraries
library(rms)
library(ggplot2)
library(dplyr)

# Define a function to generate hazard ratio data for different deciles, including confidence intervals
generate_plot_data <- function(decile_col, reference_value = 5) {
  # Convert decile column to numeric
  df_cand_kipa_w_scores[[paste0(decile_col, "_num")]] <- as.numeric(df_cand_kipa_w_scores[[decile_col]])
  
  # Ensure the datadist is set up for the model
  dd <- datadist(df_cand_kipa_w_scores)
  options(datadist = "dd")
  
  # Fit the Cox model with restricted cubic splines
  formula <- as.formula(paste("Surv(time, death_outcome) ~ rcs(", paste0(decile_col, "_num"), ", 4)"))
  model <- cph(formula, data = df_cand_kipa_w_scores, x = TRUE, y = TRUE, surv = TRUE)
  
  # Generate new data for prediction
  new_data <- data.frame(
    Decile = seq(min(df_cand_kipa_w_scores[[paste0(decile_col, "_num")]], na.rm = TRUE),
                 max(df_cand_kipa_w_scores[[paste0(decile_col, "_num")]], na.rm = TRUE),
                 length.out = 100)
  )
  names(new_data) <- paste0(decile_col, "_num")
  
  # Predict the linear predictor and standard error
  pred <- predict(model, newdata = new_data, type = "lp", se = TRUE)
  
  # Inspect the structure of the prediction object
  print("Prediction object structure:")
  print(str(pred))
  
  # Extract linear predictors and standard errors
  lp <- pred$linear.predictors
  se <- pred$se.fit
  
  # Check if lp and se are numeric
  if (is.null(lp)) stop("Linear predictor is NULL.")
  if (is.null(se)) stop("Standard error is NULL.")
  
  # Convert to numeric if necessary
  lp <- as.numeric(lp)
  se <- as.numeric(se)
  
  # Calculate hazard ratios and confidence intervals
  hr <- exp(lp)
  hr_ci_lower <- exp(lp - 1.96 * se)
  hr_ci_upper <- exp(lp + 1.96 * se)
  
  # Find the hazard ratio at the reference value (e.g., decile = 5)
  reference_hr <- hr[new_data[[paste0(decile_col, "_num")]] == reference_value]
  
  # Adjust hazard ratios so that the reference value has an HR of 1
  hr_adjusted <- hr / reference_hr
  hr_ci_lower_adjusted <- hr_ci_lower / reference_hr
  hr_ci_upper_adjusted <- hr_ci_upper / reference_hr
  
  # Create the data frame for plotting
  plot_data <- data.frame(
    Decile = new_data[[paste0(decile_col, "_num")]],
    Hazard_Ratio = hr_adjusted,
    HR_CI_Lower = hr_ci_lower_adjusted,
    HR_CI_Upper = hr_ci_upper_adjusted,
    Variable = decile_col  # Add a column for the variable name
  )
  
  return(plot_data)
}

# Generate plot data for each decile
plot_data_ADI <- generate_plot_data("ADI_decile")
plot_data_SVM <- generate_plot_data("SVM_decile")
plot_data_SVI <- generate_plot_data("SVI_decile")
plot_data_DCI <- generate_plot_data("DCI_decile")

# Combine all plot data into one data frame
combined_plot_data <- bind_rows(plot_data_ADI, plot_data_SVM, plot_data_SVI, plot_data_DCI)

# Plot all lines on the same graph with x-axis intervals of 1 and confidence intervals
ggplot(combined_plot_data, aes(x = Decile, y = Hazard_Ratio, color = Variable)) +
  geom_line() +
  geom_ribbon(aes(ymin = HR_CI_Lower, ymax = HR_CI_Upper, fill = Variable), alpha = 0.2) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Add a reference line at HR = 1
  scale_x_continuous(breaks = seq(min(combined_plot_data$Decile, na.rm = TRUE),
                                  max(combined_plot_data$Decile, na.rm = TRUE),
                                  by = 1)) +  # Set x-axis breaks to intervals of 1
  labs(title = "Hazard Ratio by Decile for Various Metrics",
       x = "Decile Number",
       y = "Hazard Ratio of Death/Deterioration (Reference = 5)",
       color = "Variable",
       fill = "Variable") +
  theme_minimal()

```
##all 4 - transplant - with confidence intervals
```{r}
# Load necessary libraries
library(rms)
library(ggplot2)
library(dplyr)

# Define a function to generate hazard ratio data for different deciles, including confidence intervals
generate_plot_data <- function(decile_col, reference_value = 5) {
  # Convert decile column to numeric
  df_cand_kipa_w_scores[[paste0(decile_col, "_num")]] <- as.numeric(df_cand_kipa_w_scores[[decile_col]])
  
  # Ensure the datadist is set up for the model
  dd <- datadist(df_cand_kipa_w_scores)
  options(datadist = "dd")
  
  # Fit the Cox model with restricted cubic splines
  formula <- as.formula(paste("Surv(time, transplant_outcome) ~ rcs(", paste0(decile_col, "_num"), ", 4)"))
  model <- cph(formula, data = df_cand_kipa_w_scores, x = TRUE, y = TRUE, surv = TRUE)
  
  # Generate new data for prediction
  new_data <- data.frame(
    Decile = seq(min(df_cand_kipa_w_scores[[paste0(decile_col, "_num")]], na.rm = TRUE),
                 max(df_cand_kipa_w_scores[[paste0(decile_col, "_num")]], na.rm = TRUE),
                 length.out = 100)
  )
  names(new_data) <- paste0(decile_col, "_num")
  
  # Predict the linear predictor and standard error
  pred <- predict(model, newdata = new_data, type = "lp", se = TRUE)
  
  # Inspect the structure of the prediction object
  print("Prediction object structure:")
  print(str(pred))
  
  # Extract linear predictors and standard errors
  lp <- pred$linear.predictors
  se <- pred$se.fit
  
  # Check if lp and se are numeric
  if (is.null(lp)) stop("Linear predictor is NULL.")
  if (is.null(se)) stop("Standard error is NULL.")
  
  # Convert to numeric if necessary
  lp <- as.numeric(lp)
  se <- as.numeric(se)
  
  # Calculate hazard ratios and confidence intervals
  hr <- exp(lp)
  hr_ci_lower <- exp(lp - 1.96 * se)
  hr_ci_upper <- exp(lp + 1.96 * se)
  
  # Find the hazard ratio at the reference value (e.g., decile = 5)
  reference_hr <- hr[new_data[[paste0(decile_col, "_num")]] == reference_value]
  
  # Adjust hazard ratios so that the reference value has an HR of 1
  hr_adjusted <- hr / reference_hr
  hr_ci_lower_adjusted <- hr_ci_lower / reference_hr
  hr_ci_upper_adjusted <- hr_ci_upper / reference_hr
  
  # Create the data frame for plotting
  plot_data <- data.frame(
    Decile = new_data[[paste0(decile_col, "_num")]],
    Hazard_Ratio = hr_adjusted,
    HR_CI_Lower = hr_ci_lower_adjusted,
    HR_CI_Upper = hr_ci_upper_adjusted,
    Variable = decile_col  # Add a column for the variable name
  )
  
  return(plot_data)
}

# Generate plot data for each decile
plot_data_ADI <- generate_plot_data("ADI_decile")
plot_data_SVM <- generate_plot_data("SVM_decile")
plot_data_SVI <- generate_plot_data("SVI_decile")
plot_data_DCI <- generate_plot_data("DCI_decile")

# Combine all plot data into one data frame
combined_plot_data <- bind_rows(plot_data_ADI, plot_data_SVM, plot_data_SVI, plot_data_DCI)

# Plot all lines on the same graph with x-axis intervals of 1 and confidence intervals
ggplot(combined_plot_data, aes(x = Decile, y = Hazard_Ratio, color = Variable)) +
  geom_line() +
  geom_ribbon(aes(ymin = HR_CI_Lower, ymax = HR_CI_Upper, fill = Variable), alpha = 0.2) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Add a reference line at HR = 1
  scale_x_continuous(breaks = seq(min(combined_plot_data$Decile, na.rm = TRUE),
                                  max(combined_plot_data$Decile, na.rm = TRUE),
                                  by = 1)) +  # Set x-axis breaks to intervals of 1
  labs(title = "Hazard Ratio by Decile for Various Metrics",
       x = "Decile Number",
       y = "Hazard Ratio of Transplant (Reference = 5)",
       color = "Variable",
       fill = "Variable") +
  theme_minimal()

```




```{r}
rm(dd, model, new_data, plot_data, plot_data_ADI, plot_data_SVM, plot_data_SVI, plot_data_DCI, ADI_seq, DCI_seq, hr, lp, SVI_seq, SVM_seq, xlim_range, ylim_range)
```
