---
title: "CIF"
date: "2024-08-09"
output:
  pdf_document: default
  html_document: default
---


```{r}
library(tidycmprsk)
library(ggplot2)
library(dplyr)
library(ggsurvfit)
```

```{r}
df_cand_kipa_w_scores$SVM_decile_num <- as.numeric(df_cand_kipa_w_scores$SVM_decile)
df_cand_kipa_w_scores$SVI_decile_num <- as.numeric(df_cand_kipa_w_scores$SVI_decile)
df_cand_kipa_w_scores$ADI_decile_num <- as.numeric(df_cand_kipa_w_scores$ADI_decile)
df_cand_kipa_w_scores$DCI_decile_num <- as.numeric(df_cand_kipa_w_scores$DCI_decile)
```

ADI - Transplant as outcome
```{r}
df_cif_transplant <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  select(PX_ID, death_outcome, time, event, ADI_decile_num)
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant <- df_cif_transplant %>% filter(!is.na(ADI_decile_num))
df_cif_transplant = df_cif_transplant %>% filter(ADI_decile_num %in% c(1,10))
```

```{r}
# Plot cuminc of transplant
cif_transplant_ADI <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ ADI_decile_num, data = df_cif_transplant) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by ADI",
    color = "ADI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_ADI
```
ADI - Death as outcome
```{r}
df_cif_death <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(death_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  select(PX_ID, death_outcome, time, event, ADI_decile_num)
```

Filter to 1st and 10th Deciles
```{r}
df_cif_death <- df_cif_death %>% filter(!is.na(ADI_decile_num))
df_cif_death = df_cif_death %>% filter(ADI_decile_num %in% c(1,10))
```

```{r}
# Plot cuminc of Death
cif_death_ADI <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ ADI_decile_num, data = df_cif_death) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Death/Deterioration",
    title = "Incidence of Death/Deterioration by ADI",
    color = "ADI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_death_ADI
```


SVM - Transplant as outcome
```{r}
df_cif_transplant_SVM <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  select(PX_ID, death_outcome, time, event, SVM_decile_num)
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant_SVM <- df_cif_transplant_SVM %>% filter(!is.na(SVM_decile_num))
df_cif_transplant_SVM = df_cif_transplant_SVM %>% filter(SVM_decile_num %in% c(1,10))
```

```{r}
# Plot cuminc of transplant
cif_transplant_SVM <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVM_decile_num, data = df_cif_transplant_SVM) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by SVM",
    color = "SVM Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVM
```

SVM - Death as outcome
```{r}
df_cif_death_SVM <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(death_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  select(PX_ID, death_outcome, time, event, SVM_decile_num)
```

Filter to 1st and 10th Deciles
```{r}
df_cif_death_SVM <- df_cif_death_SVM %>% filter(!is.na(SVM_decile_num))
df_cif_death_SVM = df_cif_death_SVM %>% filter(SVM_decile_num %in% c(1,10))
```

```{r}
# Plot cuminc of Death
cif_death_SVM <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVM_decile_num, data = df_cif_death_SVM) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Death/Deterioration",
    title = "Incidence of Death/Deterioration by SVM",
    color = "SVM Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_death_SVM
```

SVI - Transplant as outcome
```{r}
df_cif_transplant_SVI <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  select(PX_ID, death_outcome, time, event, SVI_decile_num)
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant_SVI <- df_cif_transplant_SVI %>% filter(!is.na(SVI_decile_num))
df_cif_transplant_SVI = df_cif_transplant_SVI %>% filter(SVI_decile_num %in% c(1,10))
```

```{r}
# Plot cuminc of transplant
cif_transplant_SVI <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num, data = df_cif_transplant_SVI) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by SVI",
    color = "SVI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVI
```

SVI - Death as outcome
```{r}
df_cif_death_SVI<- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(death_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  select(PX_ID, death_outcome, time, event, SVI_decile_num)
```

Filter to 1st and 10th Deciles
```{r}
df_cif_death_SVI <- df_cif_death_SVI %>% filter(!is.na(SVI_decile_num))
df_cif_death_SVI = df_cif_death_SVI %>% filter(SVI_decile_num %in% c(1,10))
```

```{r}
# Plot cuminc of Death
cif_death_SVI <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num, data = df_cif_death_SVI) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Death/Deterioration",
    title = "Incidence of Death/Deterioration by SVI",
    color = "SVI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_death_SVI
```

DCI - Transplant as outcome
```{r}
df_cif_transplant_DCI <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  select(PX_ID, death_outcome, time, event, DCI_decile_num)
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant_DCI <- df_cif_transplant_DCI %>% filter(!is.na(DCI_decile_num))
df_cif_transplant_DCI = df_cif_transplant_DCI %>% filter(DCI_decile_num %in% c(1,10))
```

```{r}
# Plot cuminc of transplant
cif_transplant_DCI <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ DCI_decile_num, data = df_cif_transplant_DCI) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by DCI",
    color = "DCI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_DCI
```

DCI - Death as outcome
```{r}
df_cif_death_DCI<- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(death_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  select(PX_ID, death_outcome, time, event, DCI_decile_num)
```

Filter to 1st and 10th Deciles
```{r}
df_cif_death_DCI <- df_cif_death_DCI %>% filter(!is.na(DCI_decile_num))
df_cif_death_DCI = df_cif_death_DCI %>% filter(DCI_decile_num %in% c(1,10))
```

```{r}
# Plot cuminc of Death
cif_death_DCI <- 
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ DCI_decile_num, data = df_cif_death_DCI) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Death/Deterioration",
    title = "Incidence of Death/Deterioration by DCI",
    color = "DCI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_death_DCI
```


```{r}
library(gridExtra)

grid.arrange(cif_transplant_SVM, cif_death_SVM, cif_transplant_SVI, cif_death_SVI, cif_transplant_ADI, cif_death_ADI, cif_transplant_DCI, cif_death_DCI, ncol = 2, nrow = 4)
```


