---
title: "SVI CIF"
author: "Olivia"
date: "2024-08-14"
output: html_document
---

```{r}
load("df_cand_kipa_w_scores_1.RData")
```


```{r eval=FALSE, include=FALSE}
df_cand_kipa_w_scores <- read.csv("df_cand_kipa_w_scores.csv")
```

```{r}
df_cand_kipa_w_scores$SVM_decile_num <- as.numeric(df_cand_kipa_w_scores$SVM_decile)
df_cand_kipa_w_scores$SVI_decile_num <- as.numeric(df_cand_kipa_w_scores$SVI_decile)
df_cand_kipa_w_scores$ADI_decile_num <- as.numeric(df_cand_kipa_w_scores$ADI_decile)
df_cand_kipa_w_scores$DCI_decile_num <- as.numeric(df_cand_kipa_w_scores$DCI_decile)
```

```{r}
df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>%
  mutate(typeO = ifelse(CAN_ABO == "O", 1, 0))
df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>%
  mutate(
    type_O  = case_when(
      typeO == 1 ~ "Type O Blood",
      typeO == 0 ~ "Other Blood Type"
  ))
```



SVI - Transplant as outcome


```{r}
df_cif_transplant_SVI <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  select(PX_ID, death_outcome, time, event, SVI_decile_num, typeO)
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant_SVI <- df_cif_transplant_SVI %>% filter(!is.na(SVI_decile_num))
df_cif_transplant_SVI = df_cif_transplant_SVI %>% filter(SVI_decile_num %in% c(1,10))
```

```{r}
library(tidycmprsk)
library(ggplot2)
library(dplyr)
library(ggsurvfit)
```


```{r}
library(tidycmprsk)
# Plot cuminc of transplant
cif_transplant_SVI <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num + typeO, data = df_cif_transplant_SVI) %>%
  ggcuminc() +
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by SVI - all dial_times",
    color = "SVI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVI
```

SVI - Transplant as outcome - Preemptive

```{r}
df_cif_transplant_SVI <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  filter(is.na(dial_time)) %>%
  select(PX_ID, death_outcome, time, event, SVI_decile_num, typeO)
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant_SVI <- df_cif_transplant_SVI %>% filter(!is.na(SVI_decile_num))
df_cif_transplant_SVI = df_cif_transplant_SVI %>% filter(SVI_decile_num %in% c(1,10))
df_cif_transplant_SVI <- df_cif_transplant_SVI %>% 
  mutate(
    SVI_decile_test = case_when(
      SVI_decile_num == 1 ~ "Decile 1",
      SVI_decile_num == 10 ~ "Decile 10"
    ))

df_cif_transplant_SVI <- df_cif_transplant_SVI %>%
  mutate(
    type_O  = case_when(
      typeO == 1 ~ "Type O Blood",
      typeO == 0 ~ "Other Blood Type"
  ))
```

```{r}
cif_transplant_SVI_preemp <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_test + type_O, data = df_cif_transplant_SVI)
cif_transplant_SVI_preemp
```


```{r}
# Plot cuminc of transplant
cif_transplant_SVI_preemp <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_test + type_O, data = df_cif_transplant_SVI) %>%
  ggcuminc()+
  labs(
    x = "Days on Waitlist",
    y = "Cumulative Incidence of Deceased Donor Transplant",
    title = "Incidence of Transplant by SVI for Preemptively Listed Patients",
    color = "SVI Decile, Blood Type"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1826)) +
  theme_minimal()+
    theme(
    # Increase font size for titles
    plot.title = element_text(size = 16),
    # Increase font size for axis labels
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    # Increase font size for axis text
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    # Increase font size for legend
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  ) +
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVI_preemp
```
SVI - Transplant as outcome - Preemptive
```{r}
df_cif_transplant_SVI <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  filter(is.na(dial_time)) %>%
  select(PX_ID, death_outcome, time, event, SVI_decile_num, typeO)

df_cif_transplant_SVI <- df_cif_transplant_SVI %>%
  filter(
    typeO == 1 & 
    (SVI_decile_num == 1 | SVI_decile_num == 10)
  )


```

Filter to 1st and 10th decile
```{r}
df_cif_transplant_SVI <- df_cif_transplant_SVI %>% filter(!is.na(SVI_decile_num))
df_cif_transplant_SVI = df_cif_transplant_SVI %>% filter(SVI_decile_num %in% c(1,10))
```

```{r}
cif_transplant_SVI_preemp <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num + typeO, data = df_cif_transplant_SVI)
cif_transplant_SVI_preemp
```



```{r}
library(gridExtra)
ggsave("SVI_decile_blood_preemp.png", plot = cif_transplant_SVI_preemp, width = 10, height = 5, dpi = 300)
```

```{r}
df_cif_transplant <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  filter(is.na(dial_time)) %>%
  select(PX_ID, transplant_outcome, time, event, SVI_decile_num, dial_time, typeO)

# Convert dial_time to numeric
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time = as.numeric(dial_time))
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant <- df_cif_transplant %>% filter(!is.na(SVI_decile_num))
df_cif_transplant = df_cif_transplant %>% filter(SVI_decile_num %in% c(1,10))
```

```{r}
cif_transplant_SVI_year <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num + typeO, data = df_cif_transplant_SVI)
cif_transplant_SVI_year
```


```{r}
# Plot cuminc of transplant
cif_transplant_SVI_year <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num + typeO, data = df_cif_transplant_SVI) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by SVI - preemptive + blood type",
    color = "SVI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVI_year
```

SVI - Transplant as outcome
```{r}
df_cif_transplant <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  filter(!is.na(dial_time)) %>%
  select(PX_ID, transplant_outcome, time, event, SVI_decile_num, dial_time, typeO)

# Convert dial_time to numeric
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time = as.numeric(dial_time))
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant <- df_cif_transplant %>% filter(!is.na(SVI_decile_num))
df_cif_transplant = df_cif_transplant %>% filter(SVI_decile_num %in% c(1,10))
```

```{r}
# Adjust breaks if the range is very small or has few unique values
max_dial_time <- max(df_cif_transplant$dial_time, na.rm = TRUE)
min_dial_time <- min(df_cif_transplant$dial_time, na.rm = TRUE)

# Ensure breaks do not exceed the range
if (max_dial_time > min_dial_time) {
  breaks <- seq(min_dial_time, max_dial_time, by = 365)
} else {
  breaks <- c(min_dial_time, max_dial_time)
}

# Create bins with adjusted breaks
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time_bin = cut(dial_time, breaks = breaks, include.lowest = TRUE))

df_cif_transplant_bin1 <- df_cif_transplant %>% 
  filter(dial_time_bin == "[0,365]")
```

```{r}
cif_transplant_SVI_year <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num + dial_time_bin + typeO, data = df_cif_transplant_bin1) 
cif_transplant_SVI_year
```


```{r}
library(cmprsk)
# Plot cuminc of transplant
cif_transplant_SVI_year <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num + dial_time_bin + typeO, data = df_cif_transplant_bin1) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by SVI - dial_time < 1 yr + blood type",
    color = "SVI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVI_year
```

SVI - Transplant as outcome
```{r}
df_cif_transplant <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  filter(!is.na(dial_time)) %>%
  select(PX_ID, transplant_outcome, time, event, SVI_decile_num, dial_time, typeO)

# Convert dial_time to numeric
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time = as.numeric(dial_time))
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant <- df_cif_transplant %>% filter(!is.na(SVI_decile_num))
df_cif_transplant = df_cif_transplant %>% filter(SVI_decile_num %in% c(1,10))
```

```{r}
# Adjust breaks if the range is very small or has few unique values
max_dial_time <- max(df_cif_transplant$dial_time, na.rm = TRUE)
min_dial_time <- min(df_cif_transplant$dial_time, na.rm = TRUE)

# Ensure breaks do not exceed the range
if (max_dial_time > min_dial_time) {
  breaks <- seq(min_dial_time, max_dial_time, by = 365)
} else {
  breaks <- c(min_dial_time, max_dial_time)
}

# Create bins with adjusted breaks
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time_bin = cut(dial_time, breaks = breaks, include.lowest = TRUE))

df_cif_transplant_bin1 <- df_cif_transplant %>% 
  filter(dial_time_bin == "(365,730]")
```

```{r}
# Plot cuminc of transplant
cif_transplant_SVI_2yrs <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num + dial_time_bin + typeO, data = df_cif_transplant_bin1) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by SVI - dial_time 1-2 yrs + blood type",
    color = "SVI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVI_2yrs
```

SVI - Transplant as outcome
```{r}
df_cif_transplant <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  filter(!is.na(dial_time)) %>%
  select(PX_ID, transplant_outcome, time, event, SVI_decile_num, dial_time, typeO)

# Convert dial_time to numeric
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time = as.numeric(dial_time))
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant <- df_cif_transplant %>% filter(!is.na(SVI_decile_num))
df_cif_transplant = df_cif_transplant %>% filter(SVI_decile_num %in% c(1,10))
```

```{r}
# Adjust breaks if the range is very small or has few unique values
max_dial_time <- max(df_cif_transplant$dial_time, na.rm = TRUE)
min_dial_time <- min(df_cif_transplant$dial_time, na.rm = TRUE)

# Ensure breaks do not exceed the range
if (max_dial_time > min_dial_time) {
  breaks <- seq(min_dial_time, max_dial_time, by = 365)
} else {
  breaks <- c(min_dial_time, max_dial_time)
}

# Create bins with adjusted breaks
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time_bin = cut(dial_time, breaks = breaks, include.lowest = TRUE))

df_cif_transplant_bin1 <- df_cif_transplant %>% 
  filter(dial_time_bin == "(730,1.09e+03]")
```

```{r}
# Plot cuminc of transplant
cif_transplant_SVI_3yrs <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num + dial_time_bin + typeO, data = df_cif_transplant_bin1) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by SVI - dial_time 2-3 yrs + blood type",
    color = "SVI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVI_3yrs
```

SVI - Transplant as outcome
```{r}
df_cif_transplant <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  filter(!is.na(dial_time)) %>%
  select(PX_ID, transplant_outcome, time, event, SVI_decile_num, dial_time, typeO)

# Convert dial_time to numeric
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time = as.numeric(dial_time))
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant <- df_cif_transplant %>% filter(!is.na(SVI_decile_num))
df_cif_transplant = df_cif_transplant %>% filter(SVI_decile_num %in% c(1,10))
```

```{r}
# Adjust breaks if the range is very small or has few unique values
max_dial_time <- max(df_cif_transplant$dial_time, na.rm = TRUE)
min_dial_time <- min(df_cif_transplant$dial_time, na.rm = TRUE)

# Ensure breaks do not exceed the range
if (max_dial_time > min_dial_time) {
  breaks <- seq(min_dial_time, max_dial_time, by = 365)
} else {
  breaks <- c(min_dial_time, max_dial_time)
}

# Create bins with adjusted breaks
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time_bin = cut(dial_time, breaks = breaks, include.lowest = TRUE))

df_cif_transplant_bin1 <- df_cif_transplant %>% 
  filter(dial_time_bin == "(1.09e+03,1.46e+03]")
```

```{r}
# Plot cuminc of transplant
cif_transplant_SVI_4yrs <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num + dial_time_bin + typeO, data = df_cif_transplant_bin1) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by SVI - dial_time 3-4 yrs + blood type",
    color = "SVI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVI_4yrs
```

SVI - Transplant as outcome - Preemptive
```{r}
df_cif_transplant_SVI <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  filter(is.na(dial_time)) %>%
  select(PX_ID, death_outcome, time, event, SVI_decile_num, typeO)
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant_SVI <- df_cif_transplant_SVI %>% filter(!is.na(SVI_decile_num))
df_cif_transplant_SVI = df_cif_transplant_SVI %>% filter(SVI_decile_num %in% c(1,10))
```

```{r}
# Plot cuminc of transplant
cif_transplant_SVI_preemp_noblood <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num, data = df_cif_transplant_SVI) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by SVI - Preemptive",
    color = "SVI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVI_preemp_noblood
```

```{r}
library(gridExtra)

SVI_dial_blood <- grid.arrange(cif_transplant_SVI, cif_transplant_SVI_preemp, cif_transplant_SVI_year, cif_transplant_SVI_2yrs, cif_transplant_SVI_3yrs, cif_transplant_SVI_4yrs, ncol = 2, nrow = 4)
```

```{r}
ggsave("SVI_dial_blood.png", plot = SVI_dial_blood, width = 20, height = 10, dpi = 300)
```


SVI - Transplant as outcome
```{r}
df_cif_transplant_SVI <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  select(PX_ID, death_outcome, time, event, SVI_decile_num, typeO)
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant_SVI <- df_cif_transplant_SVI %>% filter(!is.na(SVI_decile_num))
df_cif_transplant_SVI = df_cif_transplant_SVI %>% filter(SVI_decile_num %in% c(1,10))
```

```{r}
# Plot cuminc of transplant
cif_transplant_SVI_nb <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num, data = df_cif_transplant_SVI) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by SVI - all dial_times",
    color = "SVI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVI_nb
```

SVI - Transplant as outcome - Preemptive
```{r}
df_cif_transplant_SVI <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  filter(is.na(dial_time)) %>%
  select(PX_ID, death_outcome, time, event, SVI_decile_num, typeO)
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant_SVI <- df_cif_transplant_SVI %>% filter(!is.na(SVI_decile_num))
df_cif_transplant_SVI = df_cif_transplant_SVI %>% filter(SVI_decile_num %in% c(1,10))
```

```{r}
# Plot cuminc of transplant
cif_transplant_SVI_preemp_nb <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num, data = df_cif_transplant_SVI) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by SVI - Preemptive",
    color = "SVI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVI_preemp_nb
```

SVI - Transplant as outcome
```{r}
df_cif_transplant <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  filter(!is.na(dial_time)) %>%
  select(PX_ID, transplant_outcome, time, event, SVI_decile_num, dial_time, typeO)

# Convert dial_time to numeric
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time = as.numeric(dial_time))
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant <- df_cif_transplant %>% filter(!is.na(SVI_decile_num))
df_cif_transplant = df_cif_transplant %>% filter(SVI_decile_num %in% c(1,10))
```

```{r}
# Adjust breaks if the range is very small or has few unique values
max_dial_time <- max(df_cif_transplant$dial_time, na.rm = TRUE)
min_dial_time <- min(df_cif_transplant$dial_time, na.rm = TRUE)

# Ensure breaks do not exceed the range
if (max_dial_time > min_dial_time) {
  breaks <- seq(min_dial_time, max_dial_time, by = 365)
} else {
  breaks <- c(min_dial_time, max_dial_time)
}

# Create bins with adjusted breaks
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time_bin = cut(dial_time, breaks = breaks, include.lowest = TRUE))

df_cif_transplant_bin1 <- df_cif_transplant %>% 
  filter(dial_time_bin == "[0,365]")
```

```{r}
# Plot cuminc of transplant
cif_transplant_SVI_year_nb <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num + dial_time_bin, data = df_cif_transplant_bin1) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by SVI - dial_time < 1 yr",
    color = "SVI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVI_year_nb
```

SVI - Transplant as outcome
```{r}
df_cif_transplant <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  filter(!is.na(dial_time)) %>%
  select(PX_ID, transplant_outcome, time, event, SVI_decile_num, dial_time, typeO)

# Convert dial_time to numeric
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time = as.numeric(dial_time))
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant <- df_cif_transplant %>% filter(!is.na(SVI_decile_num))
df_cif_transplant = df_cif_transplant %>% filter(SVI_decile_num %in% c(1,10))
```

```{r}
# Adjust breaks if the range is very small or has few unique values
max_dial_time <- max(df_cif_transplant$dial_time, na.rm = TRUE)
min_dial_time <- min(df_cif_transplant$dial_time, na.rm = TRUE)

# Ensure breaks do not exceed the range
if (max_dial_time > min_dial_time) {
  breaks <- seq(min_dial_time, max_dial_time, by = 365)
} else {
  breaks <- c(min_dial_time, max_dial_time)
}

# Create bins with adjusted breaks
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time_bin = cut(dial_time, breaks = breaks, include.lowest = TRUE))

df_cif_transplant_bin1 <- df_cif_transplant %>% 
  filter(dial_time_bin == "(365,730]")
```

```{r}
# Plot cuminc of transplant
cif_transplant_SVI_2yrs_nb <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num + dial_time_bin, data = df_cif_transplant_bin1) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by SVI - dial_time 1-2 yrs",
    color = "SVI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVI_2yrs_nb
```

SVI - Transplant as outcome
```{r}
df_cif_transplant <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  filter(!is.na(dial_time)) %>%
  select(PX_ID, transplant_outcome, time, event, SVI_decile_num, dial_time, typeO)

# Convert dial_time to numeric
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time = as.numeric(dial_time))
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant <- df_cif_transplant %>% filter(!is.na(SVI_decile_num))
df_cif_transplant = df_cif_transplant %>% filter(SVI_decile_num %in% c(1,10))
```

```{r}
# Adjust breaks if the range is very small or has few unique values
max_dial_time <- max(df_cif_transplant$dial_time, na.rm = TRUE)
min_dial_time <- min(df_cif_transplant$dial_time, na.rm = TRUE)

# Ensure breaks do not exceed the range
if (max_dial_time > min_dial_time) {
  breaks <- seq(min_dial_time, max_dial_time, by = 365)
} else {
  breaks <- c(min_dial_time, max_dial_time)
}

# Create bins with adjusted breaks
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time_bin = cut(dial_time, breaks = breaks, include.lowest = TRUE))

df_cif_transplant_bin1 <- df_cif_transplant %>% 
  filter(dial_time_bin == "(730,1.09e+03]")
```

```{r}
# Plot cuminc of transplant
cif_transplant_SVI_3yrs_nb <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num + dial_time_bin, data = df_cif_transplant_bin1) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by SVI - dial_time 2-3 yrs",
    color = "SVI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVI_3yrs_nb
```

SVI - Transplant as outcome
```{r}
df_cif_transplant <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  filter(!is.na(dial_time)) %>%
  select(PX_ID, transplant_outcome, time, event, SVI_decile_num, dial_time, typeO)

# Convert dial_time to numeric
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time = as.numeric(dial_time))
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant <- df_cif_transplant %>% filter(!is.na(SVI_decile_num))
df_cif_transplant = df_cif_transplant %>% filter(SVI_decile_num %in% c(1,10))
```

```{r}
# Adjust breaks if the range is very small or has few unique values
max_dial_time <- max(df_cif_transplant$dial_time, na.rm = TRUE)
min_dial_time <- min(df_cif_transplant$dial_time, na.rm = TRUE)

# Ensure breaks do not exceed the range
if (max_dial_time > min_dial_time) {
  breaks <- seq(min_dial_time, max_dial_time, by = 365)
} else {
  breaks <- c(min_dial_time, max_dial_time)
}

# Create bins with adjusted breaks
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time_bin = cut(dial_time, breaks = breaks, include.lowest = TRUE))

df_cif_transplant_bin1 <- df_cif_transplant %>% 
  filter(dial_time_bin == "(1.09e+03,1.46e+03]")
```

```{r}
# Plot cuminc of transplant
cif_transplant_SVI_4yrs_nb <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_num + dial_time_bin, data = df_cif_transplant_bin1) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by SVI - dial_time 3-4 yrs",
    color = "SVI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVI_4yrs_nb
```

```{r}
library(gridExtra)

SVI_dial_nb <- grid.arrange(cif_transplant_SVI_nb, cif_transplant_SVI_preemp_nb, cif_transplant_SVI_year_nb, cif_transplant_SVI_2yrs_nb, cif_transplant_SVI_3yrs_nb, cif_transplant_SVI_4yrs_nb, ncol = 2, nrow = 4)
```

```{r}
ggsave("SVI_dial_nb.png", plot = SVI_dial_nb, width = 20, height = 10, dpi = 300)
```

SVI - Transplant as outcome
```{r}
df_cif_transplant_SVI <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  select(PX_ID, death_outcome, time, event, SVI_decile_num, typeO, CAN_ABO)
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant_SVI <- df_cif_transplant_SVI %>% filter(!is.na(SVI_decile_num))
df_cif_transplant_SVI = df_cif_transplant_SVI %>% filter(SVI_decile_num %in% c(1,10))
```

```{r}
# Plot cuminc of transplant
cif_transplant_SVI_test <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ typeO, data = df_cif_transplant_SVI) %>%
  ggcuminc()+
  labs(
    x = "Days",
    y = "Cumulative Incidence of Transplant",
    title = "Incidence of Transplant by SVI - all dial_times",
    color = "SVI Decile"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1800)) +
  theme_minimal()+
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVI_test
```
SVI - Transplant as outcome
```{r}
df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>%
  mutate(
    dial_time_yrs = case_when(
      dial_time < 365 ~ "0-1 year",  # Less than 1 year
      dial_time >= 365 & dial_time < 730 ~ "1-2 years",  # 1-2 years
      dial_time >= 730 & dial_time < 1095 ~ "2-3 years",  # 2-3 years
      TRUE ~ as.character(NA)  # Handle values that do not fit the criteria
    )
  )

df_cif_transplant <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  select(PX_ID, transplant_outcome, time, event, SVI_decile_num, dial_time, typeO, dial_time_yrs)

# Convert dial_time to numeric
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time = as.numeric(dial_time))
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant <- df_cif_transplant %>% filter(!is.na(SVI_decile_num))
df_cif_transplant = df_cif_transplant %>% filter(SVI_decile_num %in% c(1,10))
df_cif_transplant <- df_cif_transplant %>% 
  mutate(
    SVI_decile_test = case_when(
      SVI_decile_num == 1 ~ "Decile 1",
      SVI_decile_num == 10 ~ "Decile 10"
    ))
```

```{r}
# Adjust breaks if the range is very small or has few unique values
max_dial_time <- max(df_cif_transplant$dial_time, na.rm = TRUE)
min_dial_time <- min(df_cif_transplant$dial_time, na.rm = TRUE)

# Ensure breaks do not exceed the range
if (max_dial_time > min_dial_time) {
  breaks <- seq(min_dial_time, max_dial_time, by = 365)
} else {
  breaks <- c(min_dial_time, max_dial_time)
}

# Create bins with adjusted breaks
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time_bin = cut(dial_time, breaks = breaks, include.lowest = TRUE))
```


```{r}
df_cif_transplant_bin1 <- df_cif_transplant %>% 
  filter(dial_time_bin == "[0,365]")
```

```{r}
cif_transplant_SVI_test2 <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_test + dial_time_yrs, data = df_cif_transplant) 
cif_transplant_SVI_test2
```


```{r}
# Plot cuminc of transplant
cif_transplant_SVI_test2 <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_test + dial_time_yrs, data = df_cif_transplant) %>%
  ggcuminc()+
  labs(
    x = "Days on Waitlist",
    y = "Cumulative Incidence of Deceased Donor Transplant",
    title = "Incidence of Transplant by SVI",
    color = "SVI Decile, Dialysis Time at Listing"
    )+
  scale_ggsurvfit()+
  scale_x_continuous(limits =c(0,1826)) +
  theme_minimal()+
  theme(
    # Increase font size for titles
    plot.title = element_text(size = 16),
    # Increase font size for axis labels
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    # Increase font size for axis text
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    # Increase font size for legend
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  ) +
  guides(fill = "none", color = guide_legend(override.aes = list(linetype = 1)))
cif_transplant_SVI_test2
```
SVI - Transplant as outcome
```{r}
df_cand_kipa_w_scores <- df_cand_kipa_w_scores %>%
  mutate(
    dial_time_yrs = case_when(
      dial_time < 365 ~ "0-1 year",  # Less than 1 year
      dial_time >= 365 & dial_time < 730 ~ "1-2 years",  # 1-2 years
      dial_time >= 730 & dial_time < 1095 ~ "2-3 years",  # 2-3 years
      TRUE ~ as.character(NA)  # Handle values that do not fit the criteria
    )
  )

df_cif_transplant <- df_cand_kipa_w_scores %>%
  mutate(time = as.numeric(survival_time),  # Ensure time is numeric
         event = as.numeric(transplant_outcome)) %>%
  filter(!is.na(time) & !is.na(event)) %>%
  select(PX_ID, transplant_outcome, time, event, SVI_decile_num, dial_time, typeO, dial_time_yrs)

df_cif_transplant <- df_cif_transplant %>%
  filter(
    (SVI_decile_num == 1 & dial_time_yrs == "0-1 year") | 
    (SVI_decile_num == 10 & dial_time_yrs == "2-3 years")
  )


# Convert dial_time to numeric
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time = as.numeric(dial_time))
```

Filter to 1st and 10th decile
```{r}
df_cif_transplant <- df_cif_transplant %>% filter(!is.na(SVI_decile_num))
df_cif_transplant = df_cif_transplant %>% filter(SVI_decile_num %in% c(1,10))
df_cif_transplant <- df_cif_transplant %>% 
  mutate(
    SVI_decile_test = case_when(
      SVI_decile_num == 1 ~ "Decile 1",
      SVI_decile_num == 10 ~ "Decile 10"
    ))
```

```{r}
# Adjust breaks if the range is very small or has few unique values
max_dial_time <- max(df_cif_transplant$dial_time, na.rm = TRUE)
min_dial_time <- min(df_cif_transplant$dial_time, na.rm = TRUE)

# Ensure breaks do not exceed the range
if (max_dial_time > min_dial_time) {
  breaks <- seq(min_dial_time, max_dial_time, by = 365)
} else {
  breaks <- c(min_dial_time, max_dial_time)
}

# Create bins with adjusted breaks
df_cif_transplant <- df_cif_transplant %>%
  mutate(dial_time_bin = cut(dial_time, breaks = breaks, include.lowest = TRUE))
```

```{r}
cif_transplant_SVI_test2 <-
tidycmprsk::cuminc(Surv(time = time, event = as.factor(event)) ~ SVI_decile_test + dial_time_yrs, data = df_cif_transplant) 
cif_transplant_SVI_test2
```

```{r}
ggsave("SVI_decile_dial.png", plot = cif_transplant_SVI_test2, width = 10, height = 5, dpi = 300)
```
